---
title: "Deliveries"
subtitle: "What the coalition has sent, month by month, checked against the team's data log"
execute:
  warning: true
  message: false
---

```{r}
#| include: false
library(tidyverse)
library(here)
library(readxl)

source(here("scripts/00_utils.R"))

deliveries <- read_csv(here("data/raw/deliveries.csv"), show_col_types = FALSE)
log_path <- list.files(here("data/raw/Raw data backups"), pattern = "DATA LOG", full.names = TRUE)
```

Caring Across Generations sends data as a batch of files, roughly once a
month, and records every file in a log on Box: what it is, when it was
delivered, what dates it covers, and whether it has been checked for personal
information. The sync step reads that log, and everything on this page comes
from it. The register lists every file received; the grid below it is the
team's own month-by-month view of what has arrived, with an automatic check
that it matches the files we actually hold.

## Files received

`r nrow(deliveries)` files so far, the most recent delivered
`r format(max(deliveries$delivered, na.rm = TRUE), "%B %e, %Y")`. Search by
any word (a form number, "open", "partial", a month).

```{r}
#| echo: false
deliveries |>
  transmute(
    Delivered = delivered, File = file, Tool = tool, Form = form,
    Content = content, Status = status, Test = if_else(test, "test", ""),
    Covers = if_else(is.na(collection_start), "",
                     str_c(format(collection_start, "%b %e"), " to ", format(collection_end, "%b %e"))),
    Notes = coalesce(notes, "")
  ) |>
  arrange(desc(Delivered)) |>
  tbl_interactive(page_size = 10, searchable = TRUE, min_width = 110,
                  columns = list(File = reactable::colDef(minWidth = 320)))
```

## The team's month-by-month view

The log's Overview sheet, as the team keeps it: one row per kind of file,
one column per month, with "Received", "None", or "N/A" (not applicable yet)
in each cell.

```{r}
#| echo: false
# the sheet's first column has no header
overview <- read_excel(log_path, sheet = "Overview", col_types = "text",
                       .name_repair = "unique_quiet") |>
  rename(kind = 1) |>
  filter(!is.na(kind))
months <- names(overview)[-1]

# section banners are the rows with no month entries
overview <- overview |>
  mutate(
    banner = if_all(all_of(months), is.na),
    section = if_else(banner, kind, NA_character_)
  ) |>
  fill(section)

overview |>
  filter(!banner) |>
  select(section, kind, all_of(months)) |>
  tbl_interactive(page_size = 60, min_width = 95, groupBy = "section",
                  defaultExpanded = TRUE,
                  columns = list(section = reactable::colDef(name = "", minWidth = 170),
                                 kind = reactable::colDef(name = "", minWidth = 300)))
```

## Does the grid match the files we hold?

Each "Received" cell should correspond to a file in the register whose data
cover that month, and every file in the register should have its cell marked.
Rows the check can read are the survey and canvassing rows (a form number,
closed or open text, complete or partial); interviews and the pre-launch poll
are counted by tool. Cells listed below are worth a look rather than errors:
so far they come from a collection window logged as a single day, one logged
as "Unknown", and a file whose logged dates run into the following month.

```{r}
#| echo: false
year <- year(max(deliveries$delivered, na.rm = TRUE))

expected <- overview |>
  filter(!banner) |>
  pivot_longer(all_of(months), names_to = "month", values_to = "cell") |>
  mutate(
    k = str_to_lower(kind),
    tool = case_when(
      str_detect(k, "^[fc]\\d .*survey|^f\\d") ~ "survey",
      str_detect(k, "^c\\d")                    ~ "canvass",
      str_detect(k, "interview")                ~ "interview",
      str_detect(k, "pre-launch")               ~ "poll"
    ),
    form    = str_extract(k, "^[fc]\\d"),
    content = case_when(str_detect(k, "open") ~ "open", str_detect(k, "closed") ~ "closed"),
    status  = case_when(str_detect(k, "complete") ~ "complete", str_detect(k, "partial") ~ "partial"),
    month_start = as.Date(str_c(year, "-", match(month, month.name), "-01")),
    expected = case_when(
      is.na(cell)                                   ~ 0L,
      str_detect(str_to_lower(cell), "^n/?a$|^none$") ~ 0L,
      str_detect(cell, "\\d+")                      ~ as.integer(str_extract(cell, "\\d+")),
      str_detect(str_to_lower(cell), "received")    ~ 1L,
      TRUE                                          ~ NA_integer_
    )
  ) |>
  filter(!is.na(tool), !is.na(month_start))

# the grid's months are the months the data cover (the last collection date),
# not delivery dates; where the log has no collection dates, the delivery
# month is used. The grid's "C1" canvass row covers the earlier f1-named
# files of the same form.
actual <- deliveries |>
  filter(!test) |>
  mutate(
    form = if_else(tool == "canvass", str_replace(form, "^f", "c"), form),
    month_start = floor_date(coalesce(collection_end, delivered), "month")
  ) |>
  count(tool, form, content, status, month_start, name = "actual")

comparison <- expected |>
  left_join(actual, by = c("tool", "form", "content", "status", "month_start")) |>
  mutate(actual = replace_na(actual, 0L))

# files in the register that no readable grid row accounts for
unaccounted <- actual |>
  anti_join(expected, by = c("tool", "form", "content", "status", "month_start"))

mismatch <- comparison |> filter(expected != actual)
```

```{r}
#| echo: false
if (nrow(mismatch) == 0 && nrow(unaccounted) == 0) {
  cat("The grid and the files agree for every month checked.")
} else {
  bind_rows(
    mismatch |> transmute(Row = kind, Month = month, `Grid says` = cell, `Files held` = actual),
    unaccounted |> transmute(Row = str_c(tool, " ", coalesce(form, ""), " ", coalesce(content, ""), " ", coalesce(status, "")),
                             Month = format(month_start, "%B"), `Grid says` = "(no row or blank)", `Files held` = actual)
  ) |>
    tbl_interactive("Cells where the grid and the files differ", page_size = 60, min_width = 110)
}
```

Months the grid does not cover yet, and cells the check could not read, are
left out rather than reported as disagreements.
