---
title: "1M Care Conversations: preliminary chartbook"
author: "Manuel Alcalá Kovalski, Urban Institute"
date: today
toc-depth: 2
execute:
  warning: false
  message: false
fig-width: 7
fig-height: 4.5
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(here)
library(urbnthemes)
library(scales)
library(reactable)
library(gt)

set_urbn_defaults(style = "print")
source(here("scripts/00_utils.R"))

tbl <- function(data) {
  gt(data) |>
    sub_missing(missing_text = "—") |>
    tab_options(table.font.size = px(14), table.align = "left")
}

survey             <- read_clean("survey")
canvassing         <- read_clean("canvassing")
canvassing_revised <- read_clean("canvassing_revised")
dictionary <- read_csv(here("data/processed/data-dictionary.csv"), show_col_types = FALSE)

data_through <- max(as_date(survey$submitted_at), as_date(canvassing$submitted_at),
                    as_date(canvassing_revised$submitted_at))

# answer text for every TRUE/FALSE column, in the wording the form uses
# (the first wording where a revision added an alias)
option_labels <- dictionary |>
  filter(column %in% c(names(survey), names(canvassing))) |>
  distinct(column, .keep_all = TRUE) |>
  transmute(
    indicator = column,
    label = if_else(
      str_ends(column, "_other"),
      "Other (in their own words)",
      str_remove(option, " \\[or\\].*$")
    )
  )

# share of people shown a question who chose each answer
selection_rates <- function(data, q) {
  data |>
    pivot_selections() |>
    filter(question == q, !is.na(selected)) |>
    summarise(n_selected = sum(selected), n_shown = n(), .by = c(question, option)) |>
    mutate(indicator = str_c(question, "_", option), rate = n_selected / n_shown) |>
    left_join(option_labels, by = "indicator") |>
    mutate(label = coalesce(label, option)) |>
    filter(!(option == "other" & n_selected == 0)) # nobody wrote in: nothing to show
}

plot_rates <- function(rates, title, subtitle = NULL) {
  rates |>
    mutate(label = fct_reorder(str_wrap(label, 45), rate)) |>
    ggplot(aes(rate, label)) +
    geom_col() +
    geom_text(
      aes(label = str_glue("{percent(rate, accuracy = 1)} ({n_selected})")),
      hjust = -0.1, size = 3
    ) +
    scale_x_continuous(labels = label_percent(), expand = expansion(mult = c(0, .25))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = NULL)
}

# the same shares, split by a respondent characteristic, following the
# project's standard for subgroup reporting (Population Data Memo, July 2026):
# fewer than 50 respondents is suppressed, 50-199 is shown with a caution
# note, 200 or more is reported plainly
report_min <- 50
report_plain <- 200

rates_by <- function(data, q, group) {
  data |>
    filter(!is.na({{ group }})) |>
    pivot_selections() |>
    filter(question == q, option != "other", !is.na(selected)) |>
    summarise(rate = mean(selected), n = n(), .by = c({{ group }}, option)) |>
    filter(n >= report_min) |>
    arrange({{ group }}) |>
    mutate(
      indicator = str_c(q, "_", option),
      group_value = {{ group }},
      group = if_else(
        n < report_plain,
        str_glue("{group_value} (n = {n}; interpret with caution)"),
        str_glue("{group_value} (n = {n})")
      ),
      rate = percent(rate, accuracy = 1)
    ) |>
    left_join(option_labels, by = "indicator") |>
    select(Answer = label, group, rate) |>
    pivot_wider(names_from = group, values_from = rate)
}

share_table <- function(data, var) {
  data |>
    filter(!is.na({{ var }})) |>
    count({{ var }}, name = "respondents") |>
    mutate(share = percent(respondents / sum(respondents), accuracy = 1)) |>
    tbl()
}
```

This chartbook is a first look at who the 1M Care Conversations campaign is
reaching and what people are saying. It covers all deliveries received so far,
through **`r format(data_through, "%B %e, %Y")`**. Counts here will be lower
than the coalition's own running totals, because deliveries to Urban lag the
live forms.

Three things to keep in mind while reading:

- The three tools (the digital survey, canvassing, and interviews) are shown
  separately throughout and never added together. They reach different people
  in different ways.
- These are the people who chose to respond. The numbers describe them, not
  the country.
- Survey counts include people who stopped partway through. Each chart says
  which respondents it is based on.

The data were cleaned by `scripts/01_clean-data.qmd`, which explains how the
raw exports were handled.

## How many responses

Responses received by tool. A "partial" response is one where the person
stopped before the end; for the survey that usually means before the
demographic questions.

```{r}
#| label: counts
bind_rows(
  "survey"                    = survey,
  "canvassing"                = canvassing,
  "canvassing (revised form)" = canvassing_revised,
  .id = "tool"
) |>
  summarise(
    responses = n(),
    first = min(as_date(submitted_at)),
    last  = max(as_date(submitted_at)),
    .by = c(tool, response_status)
  ) |>
  tbl()
```

Responses by week. The survey's peaks line up with the coalition's outreach
pushes (the May 29 launch and the July 21 Virtual Day of Action).

```{r}
#| label: weekly
#| fig-height: 5
bind_rows(
  survey     = select(survey, submitted_at),
  canvassing = select(canvassing, submitted_at),
  .id = "tool"
) |>
  count(tool, week = floor_date(as_date(submitted_at), "week")) |>
  ggplot(aes(week, n)) +
  geom_col() +
  facet_wrap(~tool, ncol = 1, scales = "free_y") +
  scale_x_date(date_labels = "%b %e") +
  labs(title = "Responses received per week", x = NULL, y = "Responses")
```

## Where responses come from

The survey asks for a ZIP code at the start of the demographic section, so
we have one for `r sum(!is.na(survey$zip))` of the `r nrow(survey)` survey
respondents (the rest stopped before that point). Canvassing records the
canvasser's ZIP, not the respondent's, so it tells us where conversations
happened. Interviews are placed by the ZIP in each transcript's header; a
group session carries one ZIP for everyone in the room, and one transcript
has no ZIP at all.

```{r}
#| label: states
# USDA ERS 2020 Rural-Urban Commuting Area codes by ZIP code (Sept 2025
# release). Gives each ZIP its state and a 1-10 rurality code. Downloaded
# once into data/raw/reference/ (not committed).
ruca_path <- here("data/raw/reference/ruca2020_zip.csv")
if (!file.exists(ruca_path)) {
  dir.create(dirname(ruca_path), showWarnings = FALSE)
  download.file(
    "https://www.ers.usda.gov/media/5444/2020-rural-urban-commuting-area-codes-zip-codes.csv",
    ruca_path, headers = c("User-Agent" = "Mozilla/5.0")
  )
}
zip_lookup <- read_csv(ruca_path, col_types = cols(.default = "c")) |>
  transmute(zip = ZIPCode, state = State, ruca = as.integer(PrimaryRUCA))
zip_state <- select(zip_lookup, zip, state)

# a few people typed ZIP+4 ("97530-0524"); the first five digits are the ZIP
survey_state <- survey |>
  filter(!is.na(zip)) |>
  mutate(zip = str_sub(zip, 1, 5)) |>
  left_join(zip_lookup, by = "zip")

canvassing_state <- canvassing |>
  filter(!is.na(canvasser_zip)) |>
  left_join(zip_state, by = c("canvasser_zip" = "zip"))

# ZIP from each transcript's header (verified 2026-09-17), entered by hand
# until the coalition provides an interview metadata sheet
interviews <- tribble(
  ~transcript,                          ~format,                    ~zip,
  "nopii_grp-sf_00000_othr_20260801",   "group, short form",        "60174",
  "nopii_grp-lf_00000_othr_20260811",   "group, long form",         "49712",
  "nopii_idv-lf_00000_othr_20260818",   "individual, long form",    "60302",
  "nopii_ind-lf_00000_othr_20260819",   "individual, long form",    "48840",
  "nopii_idv-sf_00000_othr_20260821",   "individual, short form",   NA
)
interview_state <- interviews |>
  filter(!is.na(zip)) |>
  left_join(zip_state, by = "zip")

states <- tibble(
  state = c(state.abb, "DC"),
  state_name = c(state.name, "District of Columbia")
)

# adults (18+) per state from the ACS 2024 1-year, pulled once and cached
# like the national benchmarks below
acs_state_path <- here("data/raw/reference/acs2024_state_adults.csv")
if (!file.exists(acs_state_path)) {
  tidycensus::get_acs(
    "state", variables = sprintf("B01001_%03d", c(7:25, 31:49)),
    year = 2024, survey = "acs1"
  ) |>
    summarise(adults = sum(estimate), .by = NAME) |>
    rename(state_name = NAME) |>
    write_csv(acs_state_path)
}
state_adults <- read_csv(acs_state_path, show_col_types = FALSE) |>
  filter(state_name != "Puerto Rico") |>
  mutate(share_us_adults = adults / sum(adults))

# The Century Foundation's 2024 care-policy report card grade for each state
# (Gutierrez et al. 2024), as tabulated in the Population Data Memo; entered
# by hand because the report card is not available as data
care_grades <- read_csv(here("data/raw/reference/care_policy_grades_2024.csv"), show_col_types = FALSE)
failing_grades <- c("F", "D-", "D", "D+")

state_counts <- list(
  survey = survey_state,
  canvassing = canvassing_state,
  interviews = interview_state
) |>
  imap(\(data, tool) count(filter(data, !is.na(state)), state, name = tool)) |>
  reduce(full_join, by = "state") |>
  full_join(states, by = "state") |>
  mutate(
    state_name = coalesce(state_name, state),
    across(c(survey, canvassing, interviews), \(x) replace_na(x, 0L))
  ) |>
  left_join(care_grades, by = "state_name") |>
  left_join(select(state_adults, state_name, share_us_adults), by = "state_name") |>
  mutate(share_survey = survey / sum(survey)) |>
  arrange(desc(survey), state_name)

share_failing <- \(x) sum(x[state_counts$care_policy_grade %in% failing_grades]) / sum(x)
```

Every state is listed, including those with no responses yet, so gaps are
visible. Two columns give the comparison that matters for outreach: each
state's share of all survey respondents next to its share of US adults, so a
state with a much smaller survey share than population share is one the
campaign is under-reaching. The grade is the state's score on the Century
Foundation's 2024 care-policy report card, which the coalition's outreach
guidance uses to prioritize states with D and F grades. Those states hold
`r percent(share_failing(state_counts$share_us_adults), accuracy = 1)` of US
adults and `r percent(share_failing(state_counts$survey), accuracy = 1)` of
survey respondents so far. Click a column heading to sort.

```{r}
#| label: states-table
state_counts |>
  transmute(
    State = state_name, Grade = care_policy_grade,
    Survey = survey, `Share of survey` = share_survey, `Share of US adults` = share_us_adults,
    Canvassing = canvassing, Interviews = interviews
  ) |>
  reactable(
    defaultSorted = list(Survey = "desc"),
    pagination = FALSE,
    striped = TRUE,
    compact = TRUE,
    searchable = TRUE,
    defaultColDef = colDef(footer = \(values) if (is.numeric(values)) sum(values) else "Total"),
    columns = list(
      State = colDef(minWidth = 160),
      Grade = colDef(maxWidth = 70, footer = ""),
      `Share of survey` = colDef(format = colFormat(percent = TRUE, digits = 1), footer = \(v) percent(sum(v), accuracy = 1)),
      `Share of US adults` = colDef(format = colFormat(percent = TRUE, digits = 1), footer = \(v) percent(sum(v), accuracy = 1))
    )
  )
```

`r sum(is.na(survey_state$state))` survey ZIP codes and
`r sum(is.na(canvassing_state$state))` canvasser ZIP codes did not match a
real ZIP (for example, `00000`) and are left out of the table.

```{r}
#| label: states-chart
#| fig-height: 5
state_counts |>
  filter(survey > 0) |>
  slice_max(survey, n = 15) |>
  mutate(state_name = fct_reorder(state_name, survey)) |>
  ggplot(aes(survey, state_name)) +
  geom_col() +
  geom_text(aes(label = survey), hjust = -0.3, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(
    title = "Survey respondents by state, top 15",
    subtitle = str_glue("Among {nrow(survey_state)} respondents who gave a ZIP code"),
    x = "Respondents", y = NULL
  )
```

### Rural and urban

The same ZIP lookup classifies each ZIP code by how rural it is, using the
[Rural-Urban Commuting Area (RUCA) codes](https://www.ers.usda.gov/data-products/rural-urban-commuting-area-codes)
from the USDA Economic Research Service. The codes run from 1 (inside a
metropolitan area) to 10 (rural, far from any town), based on where people in
that ZIP commute to work. We group them two ways: a four-way split used in
much rural health research, and the urban/rural split the coalition asked for
(Jaimie Worker, CAG, September 2026), where "rural" is RUCA 4 to 10. That is
also the cut the Health Resources and Services Administration's Federal
Office of Rural Health Policy uses to
[define rural](https://www.hrsa.gov/rural-health/about-us/what-is-rural) for
its programs.

```{r}
#| label: ruca
survey_state <- survey_state |>
  mutate(
    rurality = case_when(
      ruca %in% 1:3  ~ "Metropolitan (1-3)",
      ruca %in% 4:6  ~ "Micropolitan (4-6)",
      ruca %in% 7:9  ~ "Small town (7-9)",
      ruca == 10     ~ "Rural (10)"
    ) |> factor(levels = c("Metropolitan (1-3)", "Micropolitan (4-6)", "Small town (7-9)", "Rural (10)")),
    urban_rural = if_else(ruca %in% 1:3, "Urban (RUCA 1-3)", "Rural (RUCA 4-10)") |>
      factor(levels = c("Urban (RUCA 1-3)", "Rural (RUCA 4-10)"))
  )

survey_state |>
  filter(!is.na(rurality)) |>
  count(urban_rural, rurality, name = "respondents") |>
  mutate(share = percent(respondents / sum(respondents), accuracy = 1)) |>
  rename("Urban or rural" = urban_rural, "RUCA group" = rurality) |>
  tbl()
```

Rural respondents are a smaller group, so splits within the rural categories
are still thin. The urban/rural split is large enough to compare answers
under the project's subgroup standard (200 or more respondents per group).
For example, what ideal care looks like:

```{r}
#| label: q6-by-rurality
rates_by(survey_state, "q6", urban_rural) |>
  tbl()
```

One check on that definition. RUCA codes 2 and 3 are places within
commuting range of a metropolitan core but outside it — suburbs and
exurbs — and whether they count as urban or rural is a judgment call. The
same question split three ways shows whether that choice matters: if the
middle column looks like the first, the definition above is safe; if it looks
like the last, counting 2 and 3 as rural would change the picture.

```{r}
#| label: q6-by-ruca3
survey_state |>
  mutate(ruca_group = case_when(
    ruca == 1        ~ "RUCA 1 (metro core)",
    ruca %in% 2:3    ~ "RUCA 2-3 (metro commuting)",
    ruca >= 4        ~ "RUCA 4-10 (rural)"
  ) |> factor(levels = c("RUCA 1 (metro core)", "RUCA 2-3 (metro commuting)", "RUCA 4-10 (rural)"))) |>
  rates_by("q6", ruca_group) |>
  tbl()
```

## Who is responding to the survey

```{r}
#| label: demographics-data
#| include: false
demographics <- survey |>
  filter(if_any(c(zip, hh_size, hh_income, age, disability, political), \(x) !is.na(x)))
```

Only the survey asks demographic questions; canvassing and interviews do not.
The tables in this section cover the `r nrow(demographics)` survey
respondents who reached the demographic section. Shares are among people who
answered each question, and "Prefer not to say" is counted as an answer.

### Age

Respondents skew older: most are 55 or over.

```{r}
#| label: age
demographics |>
  filter(!is.na(age)) |>
  count(age) |>
  ggplot(aes(n, fct_rev(age))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.3, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(title = "Age", x = "Respondents", y = NULL)
```

### Race and ethnicity

People could choose more than one answer, so shares add to more than 100
percent.

```{r}
#| label: race
demographics |>
  selection_rates("race") |>
  plot_rates("Race and ethnicity", "Share of respondents choosing each; more than one allowed")
```

### Gender

```{r}
#| label: gender
demographics |>
  selection_rates("gender") |>
  transmute(gender = label, respondents = n_selected, share = percent(rate, accuracy = 1)) |>
  arrange(desc(respondents)) |>
  tbl()
```

### Household income

```{r}
#| label: income
share_table(demographics, hh_income)
```

### Household size

```{r}
#| label: hh-size
share_table(demographics, hh_size)
```

### Disability or chronic health condition

```{r}
#| label: disability
share_table(demographics, disability)
```

### Political views

```{r}
#| label: political
share_table(demographics, political)
```

### Compared with US adults

To see who the campaign is reaching and missing, each characteristic is set
next to the same figure for the country, from the Census Bureau's American
Community Survey (2024 one-year estimates). This is context, not a claim that
the survey resembles the population: respondents chose to take part, and the
gaps are the point. "Prefer not to say" answers are left out of the survey
shares here so the two columns add up the same way.

Some survey categories do not line up with the Census exactly; each table
says how they were matched. Disability and political views are not compared:
the Census defines disability differently, and there is no Census measure of
political views.

```{r}
#| label: acs-pull
#| include: false
# ACS 2024 1-year, national. Pulled once with tidycensus (needs CENSUS_API_KEY)
# and kept in data/raw/reference/ so rendering works offline afterwards.
acs_path <- here("data/raw/reference/acs2024_benchmarks.csv")
if (!file.exists(acs_path)) {
  acs_vars <- c(
    sprintf("B01001_%03d", c(7:25, 31:49)),   # sex by age, adults
    sprintf("B03002_%03d", c(1, 3:9, 12)),    # Hispanic origin by race
    sprintf("B19001_%03d", 1:17),             # household income
    sprintf("B11016_%03d", c(3:8, 10:16))     # household size
  )
  tidycensus::get_acs("us", variables = acs_vars, year = 2024, survey = "acs1") |>
    select(variable, estimate) |>
    write_csv(acs_path)
}
acs <- read_csv(acs_path, show_col_types = FALSE) |> deframe()
acs_sum <- \(table, idx) sum(acs[sprintf("%s_%03d", table, idx)])

# one comparison table: category, survey share, US share
compare_table <- function(survey_shares, us_counts, title, note, us_total = sum(us_counts$us, na.rm = TRUE)) {
  survey_shares |>
    full_join(us_counts |> mutate(us = us / us_total), by = "category") |>
    mutate(category = factor(category, levels = us_counts$category)) |> # keep the questionnaire's order
    arrange(category) |>
    gt() |>
    cols_label(category = "", survey = "Survey", us = "US adults (ACS 2024)") |>
    fmt_percent(c(survey, us), decimals = 0) |>
    sub_missing(missing_text = "—") |>
    tab_header(title = title) |>
    tab_source_note(note) |>
    opt_align_table_header("left") |>
    tab_options(table.font.size = px(14), table.align = "left")
}

# share of respondents in each level of a single-answer question, dropping
# "Prefer not to say"
survey_shares <- function(data, var, recode = identity) {
  data |>
    filter(!is.na({{ var }}), {{ var }} != "Prefer not to say") |>
    mutate(category = recode(as.character({{ var }}))) |>
    count(category) |>
    mutate(survey = n / sum(n)) |>
    select(category, survey)
}
```

```{r}
#| label: acs-age
compare_table(
  survey_shares(demographics, age),
  tribble(
    ~category,     ~us,
    "18-24",       acs_sum("B01001", c(7:10, 31:34)),
    "25-34",       acs_sum("B01001", c(11:12, 35:36)),
    "35-44",       acs_sum("B01001", c(13:14, 37:38)),
    "45-54",       acs_sum("B01001", c(15:16, 39:40)),
    "55-64",       acs_sum("B01001", c(17:19, 41:43)),
    "65-74",       acs_sum("B01001", c(20:22, 44:46)),
    "75 or older", acs_sum("B01001", c(23:25, 47:49))
  ),
  "Age",
  "US figure: population 18 and older."
)
```

```{r}
#| label: acs-gender
gender_shares <- demographics |>
  filter(!is.na(gender_woman), !gender_prefer_not_to_say) |>
  summarise(Woman = mean(gender_woman), Man = mean(gender_man)) |>
  pivot_longer(everything(), names_to = "category", values_to = "survey")

compare_table(
  gender_shares,
  tribble(
    ~category, ~us,
    "Woman",   acs_sum("B01001", 31:49),
    "Man",     acs_sum("B01001", 7:25)
  ),
  "Gender",
  "Survey: share of respondents choosing each (more than one allowed; non-binary and transgender have no Census counterpart). US figure: female and male population 18 and older."
)
```

```{r}
#| label: acs-race
race_shares <- demographics |>
  filter(!is.na(race_white), !race_prefer_not_to_say) |>
  summarise(
    "Hispanic, Latino, or Latina"       = mean(race_hispanic),
    "White"                             = mean(race_white),
    "Black or African American"         = mean(race_black),
    "Asian or Asian American"           = mean(race_asian),
    "American Indian or Alaska Native"  = mean(race_aian),
    "Native Hawaiian or Pacific Islander" = mean(race_nhpi),
    "Multiracial"                       = mean(race_multiracial),
    "Middle Eastern or North African"   = mean(race_mena)
  ) |>
  pivot_longer(everything(), names_to = "category", values_to = "survey")

compare_table(
  race_shares,
  tribble(
    ~category,                            ~us,
    "Hispanic, Latino, or Latina",        acs["B03002_012"],
    "White",                              acs["B03002_003"],
    "Black or African American",          acs["B03002_004"],
    "Asian or Asian American",            acs["B03002_006"],
    "American Indian or Alaska Native",   acs["B03002_005"],
    "Native Hawaiian or Pacific Islander", acs["B03002_007"],
    "Multiracial",                        acs["B03002_009"],
    "Middle Eastern or North African",    NA
  ),
  "Race and ethnicity",
  "Survey: share choosing each (more than one allowed, so shares exceed 100%). US figure: Hispanic of any race, then non-Hispanic single races and two-or-more races, as a share of the total population; the Census does not yet report Middle Eastern or North African separately. Non-Hispanic 'some other race' (about 1%) is not shown.",
  us_total = acs["B03002_001"]
)
```

```{r}
#| label: acs-income
income_to_acs <- \(x) case_match(
  x,
  "Under $25,000" ~ "Under $25,000",
  "$25,000-$49,999" ~ "$25,000-$49,999",
  "$50,000-$74,999" ~ "$50,000-$74,999",
  "$75,000-$99,999" ~ "$75,000-$99,999",
  "$100,000-$149,999" ~ "$100,000-$149,999",
  c("$150,000-$174,999", "$175,000-$199,999") ~ "$150,000-$199,999",
  c("$200,000-$249,999", "$250,000-$299,999", "$300,000 or more") ~ "$200,000 or more"
)

compare_table(
  survey_shares(demographics, hh_income, income_to_acs),
  tribble(
    ~category,           ~us,
    "Under $25,000",     acs_sum("B19001", 2:5),
    "$25,000-$49,999",   acs_sum("B19001", 6:10),
    "$50,000-$74,999",   acs_sum("B19001", 11:12),
    "$75,000-$99,999",   acs_sum("B19001", 13),
    "$100,000-$149,999", acs_sum("B19001", 14:15),
    "$150,000-$199,999", acs_sum("B19001", 16),
    "$200,000 or more",  acs_sum("B19001", 17)
  ),
  "Household income",
  "Survey brackets above $150,000 are combined to match the Census. Survey: respondents; US figure: households."
)
```

```{r}
#| label: acs-hh-size
compare_table(
  survey_shares(demographics, hh_size, \(x) if_else(x == "1, just myself", "1", x)),
  tribble(
    ~category,   ~us,
    "1",         acs_sum("B11016", 10),
    "2",         acs_sum("B11016", c(3, 11)),
    "3",         acs_sum("B11016", c(4, 12)),
    "4",         acs_sum("B11016", c(5, 13)),
    "5",         acs_sum("B11016", c(6, 14)),
    "6",         acs_sum("B11016", c(7, 15)),
    "7 or more", acs_sum("B11016", c(8, 16))
  ),
  "Household size",
  "Survey: respondents; US figure: households. A person-based US figure would tilt toward larger households."
)
```

## Connection to care

The first question asks about a person's connection to care and routes them
to one of five pathways. This decides which questions each person sees, so it
shapes everything below. Current caregivers and care recipients are by far
the largest group.

```{r}
#| label: pathways
bind_rows(survey = survey, canvassing = canvassing, .id = "tool") |>
  filter(!is.na(pathway)) |>
  count(tool, pathway) |>
  mutate(share = percent(n / sum(n), accuracy = 1), .by = tool) |>
  pivot_wider(names_from = tool, values_from = c(n, share)) |>
  select(pathway, survey_n = n_survey, survey_share = share_survey,
         canvassing_n = n_canvassing, canvassing_share = share_canvassing) |>
  tbl()
```

The answers behind the pathways. People could choose more than one.

```{r}
#| label: q1
#| fig-height: 5
survey |>
  selection_rates("q1") |>
  plot_rates(
    "Connection to care (survey)",
    str_glue("Share of {nrow(filter(survey, !is.na(q1_child_now)))} respondents who answered; more than one allowed")
  )
```

## What people are saying

For each question, the chart shows the share of people *who were shown the
question* choosing each answer, with the number of people in parentheses.
Because the pathways see different questions, the group behind each chart is
different; the subtitle says who.

### Challenges for current caregivers and care recipients

```{r}
#| label: q2a
#| fig-height: 5
q2a <- selection_rates(survey, "q2a")
plot_rates(
  q2a,
  "What challenges have you experienced with care?",
  str_glue("Survey, current pathway, {unique(q2a$n_shown)} respondents")
)
```

### Supports that would make a difference for current caregivers and recipients

```{r}
#| label: q3a
#| fig-height: 5
q3a <- selection_rates(survey, "q3a")
plot_rates(
  q3a,
  "What care support would make a difference for you?",
  str_glue("Survey, current pathway, {unique(q3a$n_shown)} respondents")
)
```

### Who helps most with care

```{r}
#| label: q4a
#| fig-height: 5
q4a <- selection_rates(survey, "q4a")
plot_rates(
  q4a,
  "What or who helps you most with care? (top 3)",
  str_glue("Survey, current pathway, {unique(q4a$n_shown)} respondents")
)
```

### What ideal care would look like

Everyone sees this question regardless of pathway, so it has the widest base.
It asks people to choose up to three answers, but the form does not enforce
that, and roughly four in ten respondents chose more — some chose all eleven.
The table shows shares two ways: among everyone, and among only the people
who kept to three or fewer. The ranking is similar either way, but the
levels differ, and it is worth deciding which version to report.

```{r}
#| label: q6
#| fig-height: 5.5
q6_all <- selection_rates(survey, "q6")

q6_n_chosen <- survey |>
  filter(!is.na(q6_affordable)) |>
  mutate(n_chosen = rowSums(across(starts_with("q6_") & where(is.logical) & !q6_other))) |>
  select(respondent_id, n_chosen)

q6_within <- survey |>
  semi_join(filter(q6_n_chosen, n_chosen <= 3), by = "respondent_id") |>
  selection_rates("q6")

plot_rates(
  q6_all,
  "What would ideal care support and services look like for you?",
  str_glue("Survey, all pathways, {unique(q6_all$n_shown)} respondents")
)
```

```{r}
#| label: q6-table
q6_all |>
  select(label, everyone = rate) |>
  left_join(select(q6_within, label, three_or_fewer = rate), by = "label") |>
  arrange(desc(everyone)) |>
  mutate(across(c(everyone, three_or_fewer), \(x) percent(x, accuracy = 1))) |>
  rename(
    Answer = label,
    "Everyone ({unique(q6_all$n_shown)})" := everyone,
    "Chose 3 or fewer ({unique(q6_within$n_shown)})" := three_or_fewer
  ) |>
  tbl()
```

### Ideal care by respondent group

The same question, split by age and by disability status. Each column is one
group; the number in the heading is how many people in that group answered.
These splits follow the project's standard for reporting on subgroups: a
group with fewer than 50 respondents is not shown, a group with 50 to 199 is
marked "interpret with caution," and a group with 200 or more is reported
plainly. With more responses these splits can extend to every question and
every demographic.

```{r}
#| label: q6-by-age
rates_by(survey, "q6", age) |>
  tbl()
```

```{r}
#| label: q6-by-disability
rates_by(survey, "q6", disability) |>
  tbl()
```

### In their own words

Every choose-all-that-apply question has an "Other" box. Text written there
is kept word for word and will be coded alongside the open-ended questions.
Number of respondents who wrote something, by question:

```{r}
#| label: other-text
survey |>
  select(ends_with("_other_text")) |>
  summarise(across(everything(), \(x) sum(!is.na(x)))) |>
  pivot_longer(everything(), names_to = "question", values_to = "respondents") |>
  mutate(question = str_remove(question, "_other_text")) |>
  filter(respondents > 0) |>
  arrange(desc(respondents)) |>
  tbl()
```

## Canvassing

Canvassing is a doorstep conversation, so it is a funnel: not everyone
approached answers, not everyone who answers has a connection to care, and
not everyone with a connection agrees to the full questions. Each stage is a
different base for any analysis.

```{r}
#| label: funnel
funnel <- tibble(
  stage = c(
    "Approached",
    "Answered and engaged",
    "Reports a connection to care",
    "Agreed to full questions"
  ),
  n = c(
    nrow(canvassing),
    sum(canvassing$responded == "Yes - continue with questions", na.rm = TRUE),
    sum(canvassing$care_connection == "Yes", na.rm = TRUE),
    sum(!is.na(canvassing$pathway))
  )
) |>
  mutate(stage = fct_rev(fct_inorder(stage)))

ggplot(funnel, aes(n, stage)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.5, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(title = "Canvassing doorstep funnel", x = "Contacts", y = NULL)
```

At the door, the canvasser records challenges from a short list before the
full questions begin. This covers everyone who engaged, not just those who
went on to the full questions.

```{r}
#| label: doorstep
doorstep <- selection_rates(canvassing, "doorstep")
plot_rates(
  doorstep,
  "Challenges mentioned at the door",
  str_glue("Canvassing, {unique(doorstep$n_shown)} engaged contacts")
)
```

The full question set was answered by too few canvassing contacts to chart
yet (`r sum(!is.na(canvassing$q1_child_now))` so far). Those answers are in
the clean file and will be added here as canvassing grows.

### The revised form (August on)

In August the coalition changed the canvassing form: the yes/no
connection-to-care question and the challenge checklist were replaced by
"Do you need care or is there someone in your life you help take care of?"
and a free-text "What's been hard about that?" There are three versions of
the revised form, with different questions; almost all responses so far come
from the Daisy Chain event (`c2`), which asked only the two new questions.
Because the questions differ, these contacts are kept apart from the funnel
and challenge charts above.

```{r}
#| label: canvassing-revised
canvassing_revised |>
  count(form_version, response_status, name = "contacts") |>
  rename("Form version" = form_version, "Status" = response_status) |>
  tbl()
```

```{r}
#| label: canvassing-revised-screener
canvassing_revised |>
  mutate(screener = fct_na_value_to_level(screener, "Not answered")) |>
  count(screener, wrote_text = !is.na(hard_text), name = "contacts") |>
  rename("Needs care or helps care for someone" = screener,
         "Wrote something for \"What's been hard\"" = wrote_text) |>
  tbl()
```

The `r sum(!is.na(canvassing_revised$hard_text))` free-text answers are
respondents' own words and go to the text analysis with the open-ended
survey questions.

## Interviews

Five interview transcripts have been delivered: two group sessions and three
individual interviews, in Illinois and Michigan (one transcript carries no
ZIP code). Interview content is being coded against the project codebook and
reported in the coding dashboard.

```{r}
#| label: interviews
interviews |>
  left_join(zip_state, by = "zip") |>
  left_join(states, by = "state") |>
  select(transcript, format, zip, state = state_name) |>
  tbl()
```

## Not in this chartbook yet

- **Rural detail.** Rural respondents can be split further (small towns,
  remote areas, commuting patterns) once there are enough of them.
- **Open-ended answers.** The survey's and canvassing's open-text questions
  are coded separately once the codebook is updated (issue #2).
- **Later channels.** The email opt-in, Spanish, and film-screening survey
  links and the second canvassing event have not been delivered yet.
