---
title: "1M Care Conversations: preliminary chartbook"
author: "Manuel Alcalá Kovalski, Urban Institute"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
execute:
  warning: false
  message: false
fig-width: 7
fig-height: 4.5
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(here)
library(urbnthemes)
library(scales)
library(reactable)

set_urbn_defaults(style = "print")
source(here("scripts/00_utils.R"))

survey     <- read_rds(here("data/processed/survey_clean.rds"))
canvassing <- read_rds(here("data/processed/canvassing_clean.rds"))
dictionary <- read_csv(here("data/processed/data-dictionary.csv"), show_col_types = FALSE)

data_through <- max(as_date(survey$submitted_at), as_date(canvassing$submitted_at))

# answer text for every TRUE/FALSE column, in the wording the form uses
# (the first wording where a revision added an alias)
option_labels <- dictionary |>
  filter(column %in% c(names(survey), names(canvassing))) |>
  distinct(column, .keep_all = TRUE) |>
  transmute(
    indicator = column,
    label = if_else(
      str_ends(column, "_other"),
      "Other (in their own words)",
      str_remove(option, " \\[or\\].*$")
    )
  )

# share of people shown a question who chose each answer
selection_rates <- function(data, q) {
  data |>
    pivot_selections() |>
    filter(question == q, !is.na(selected)) |>
    summarise(n_selected = sum(selected), n_shown = n(), .by = c(question, option)) |>
    mutate(indicator = str_c(question, "_", option), rate = n_selected / n_shown) |>
    left_join(option_labels, by = "indicator") |>
    mutate(label = coalesce(label, option)) |>
    filter(!(option == "other" & n_selected == 0)) # nobody wrote in: nothing to show
}

plot_rates <- function(rates, title, subtitle = NULL) {
  rates |>
    mutate(label = fct_reorder(str_wrap(label, 45), rate)) |>
    ggplot(aes(rate, label)) +
    geom_col() +
    geom_text(
      aes(label = str_glue("{percent(rate, accuracy = 1)} ({n_selected})")),
      hjust = -0.1, size = 3
    ) +
    scale_x_continuous(labels = label_percent(), expand = expansion(mult = c(0, .25))) +
    labs(title = title, subtitle = subtitle, x = NULL, y = NULL)
}

# the same shares, split by a respondent characteristic; groups with fewer
# than `min_n` respondents are left out rather than shown on thin evidence
rates_by <- function(data, q, group, min_n = 30) {
  data |>
    filter(!is.na({{ group }})) |>
    pivot_selections() |>
    filter(question == q, option != "other", !is.na(selected)) |>
    summarise(rate = mean(selected), n = n(), .by = c({{ group }}, option)) |>
    filter(n >= min_n) |>
    arrange({{ group }}) |>
    mutate(
      indicator = str_c(q, "_", option),
      group_value = {{ group }},
      group = str_glue("{group_value} (n = {n})"),
      rate = percent(rate, accuracy = 1)
    ) |>
    left_join(option_labels, by = "indicator") |>
    select(Answer = label, group, rate) |>
    pivot_wider(names_from = group, values_from = rate)
}

share_table <- function(data, var) {
  data |>
    filter(!is.na({{ var }})) |>
    count({{ var }}, name = "respondents") |>
    mutate(share = percent(respondents / sum(respondents), accuracy = 1)) |>
    knitr::kable()
}
```

This chartbook is a first look at who the 1M Care Conversations campaign is
reaching and what people are saying. It covers all deliveries received so far,
through **`r format(data_through, "%B %e, %Y")`**. Counts here will be lower
than the coalition's own running totals, because deliveries to Urban lag the
live forms.

Three things to keep in mind while reading:

- The three tools — the digital survey, canvassing, and interviews — are shown
  separately throughout and never added together. They reach different people
  in different ways.
- These are the people who chose to respond. The numbers describe them, not
  the country; nothing here should be read as nationally representative.
- Survey counts include people who stopped partway through. Each chart says
  which respondents it is based on.

The data were cleaned by `scripts/01_clean-data.qmd`, which explains how the
raw exports were handled.

## How many responses

Responses received by tool. A "partial" response is one where the person
stopped before the end; for the survey that usually means before the
demographic questions.

```{r}
#| label: counts
bind_rows(
  survey     = survey,
  canvassing = canvassing,
  .id = "tool"
) |>
  summarise(
    responses = n(),
    first = min(as_date(submitted_at)),
    last  = max(as_date(submitted_at)),
    .by = c(tool, response_status)
  ) |>
  knitr::kable()
```

Responses by week. The survey's peaks line up with the coalition's outreach
pushes (the May 29 launch and the July 21 Virtual Day of Action).

```{r}
#| label: weekly
#| fig-height: 5
bind_rows(
  survey     = select(survey, submitted_at),
  canvassing = select(canvassing, submitted_at),
  .id = "tool"
) |>
  count(tool, week = floor_date(as_date(submitted_at), "week")) |>
  ggplot(aes(week, n)) +
  geom_col() +
  facet_wrap(~tool, ncol = 1, scales = "free_y") +
  scale_x_date(date_labels = "%b %e") +
  labs(title = "Responses received per week", x = NULL, y = "Responses")
```

## Where responses come from

The survey asks for a ZIP code at the start of the demographic section, so
we have one for `r sum(!is.na(survey$zip))` of the `r nrow(survey)` survey
respondents (the rest stopped before that point). Canvassing records the
canvasser's ZIP, not the respondent's, so it tells us where conversations
happened. Interviews are placed by the ZIP in each transcript's header; a
group session carries one ZIP for everyone in the room, and one transcript
has no ZIP at all.

```{r}
#| label: states
# USDA ERS 2020 Rural-Urban Commuting Area codes by ZIP code (Sept 2025
# release). Gives each ZIP its state and a 1-10 rurality code. Downloaded
# once into data/raw/reference/ (not committed).
ruca_path <- here("data/raw/reference/ruca2020_zip.csv")
if (!file.exists(ruca_path)) {
  dir.create(dirname(ruca_path), showWarnings = FALSE)
  download.file(
    "https://www.ers.usda.gov/media/5444/2020-rural-urban-commuting-area-codes-zip-codes.csv",
    ruca_path, headers = c("User-Agent" = "Mozilla/5.0")
  )
}
zip_lookup <- read_csv(ruca_path, col_types = cols(.default = "c")) |>
  transmute(zip = ZIPCode, state = State, ruca = as.integer(PrimaryRUCA))
zip_state <- select(zip_lookup, zip, state)

# a few people typed ZIP+4 ("97530-0524"); the first five digits are the ZIP
survey_state <- survey |>
  filter(!is.na(zip)) |>
  mutate(zip = str_sub(zip, 1, 5)) |>
  left_join(zip_lookup, by = "zip")

canvassing_state <- canvassing |>
  filter(!is.na(canvasser_zip)) |>
  left_join(zip_state, by = c("canvasser_zip" = "zip"))

# ZIP from each transcript's header (verified 2026-09-17), entered by hand
# until the coalition provides an interview metadata sheet
interviews <- tribble(
  ~transcript,                          ~format,                  ~zip,
  "nopii_grp-sf_00000_othr_20260801",   "group session",          "60174",
  "nopii_idv-lf_00000_othr_20260818",   "individual, long form",  "60302",
  "nopii_idv-sf_00000_othr_20260821",   "individual, short form", NA
)
interview_state <- interviews |>
  filter(!is.na(zip)) |>
  left_join(zip_state, by = "zip")

states <- tibble(
  state = c(state.abb, "DC"),
  state_name = c(state.name, "District of Columbia")
)

state_counts <- list(
  survey = survey_state,
  canvassing = canvassing_state,
  interviews = interview_state
) |>
  imap(\(data, tool) count(filter(data, !is.na(state)), state, name = tool)) |>
  reduce(full_join, by = "state") |>
  full_join(states, by = "state") |>
  mutate(
    state_name = coalesce(state_name, state),
    across(c(survey, canvassing, interviews), \(x) replace_na(x, 0L))
  ) |>
  arrange(desc(survey), state_name)
```

Every state is listed, including those with no responses yet, so gaps are
visible. Click a column heading to sort.

```{r}
#| label: states-table
state_counts |>
  select(State = state_name, Survey = survey, Canvassing = canvassing, Interviews = interviews) |>
  reactable(
    defaultSorted = list(Survey = "desc"),
    pagination = FALSE,
    striped = TRUE,
    compact = TRUE,
    searchable = TRUE,
    defaultColDef = colDef(footer = \(values) if (is.numeric(values)) sum(values) else "Total"),
    columns = list(State = colDef(minWidth = 160))
  )
```

`r sum(is.na(survey_state$state))` survey ZIP codes and
`r sum(is.na(canvassing_state$state))` canvasser ZIP codes did not match a
real ZIP (for example, `00000`) and are left out of the table.

```{r}
#| label: states-chart
#| fig-height: 5
state_counts |>
  filter(survey > 0) |>
  slice_max(survey, n = 15) |>
  mutate(state_name = fct_reorder(state_name, survey)) |>
  ggplot(aes(survey, state_name)) +
  geom_col() +
  geom_text(aes(label = survey), hjust = -0.3, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(
    title = "Survey respondents by state, top 15",
    subtitle = str_glue("Among {nrow(survey_state)} respondents who gave a ZIP code"),
    x = "Respondents", y = NULL
  )
```

### Rural and urban

The same ZIP lookup classifies each ZIP code by how rural it is, using the
USDA's Rural-Urban Commuting Area (RUCA) codes. The codes run from 1 (inside
a metropolitan area) to 10 (rural, far from any town), based on where people
in that ZIP commute to work. We group them two ways: a four-way split used in
much rural health research, and the simple urban/rural split the coalition
asked about, where "rural" is everything outside metropolitan areas (RUCA 4
to 10).

```{r}
#| label: ruca
survey_state <- survey_state |>
  mutate(
    rurality = case_when(
      ruca %in% 1:3  ~ "Metropolitan (1-3)",
      ruca %in% 4:6  ~ "Micropolitan (4-6)",
      ruca %in% 7:9  ~ "Small town (7-9)",
      ruca == 10     ~ "Rural (10)"
    ) |> factor(levels = c("Metropolitan (1-3)", "Micropolitan (4-6)", "Small town (7-9)", "Rural (10)")),
    urban_rural = if_else(ruca %in% 1:3, "Urban (RUCA 1-3)", "Rural (RUCA 4-10)") |>
      factor(levels = c("Urban (RUCA 1-3)", "Rural (RUCA 4-10)"))
  )

survey_state |>
  filter(!is.na(rurality)) |>
  count(urban_rural, rurality, name = "respondents") |>
  mutate(share = percent(respondents / sum(respondents), accuracy = 1)) |>
  rename("Urban or rural" = urban_rural, "RUCA group" = rurality) |>
  knitr::kable()
```

Rural respondents are a small group so far, so splits within the rural
categories are not yet reliable. The urban/rural split, though, is large
enough to compare answers — for example, what ideal care looks like:

```{r}
#| label: q6-by-rurality
rates_by(survey_state, "q6", urban_rural) |>
  knitr::kable()
```

## Who is responding to the survey

```{r}
#| label: demographics-data
#| include: false
demographics <- survey |>
  filter(if_any(c(zip, hh_size, hh_income, age, disability, political), \(x) !is.na(x)))
```

Only the survey asks demographic questions; canvassing and interviews do not.
The tables in this section cover the `r nrow(demographics)` survey
respondents who reached the demographic section. Shares are among people who
answered each question, and "Prefer not to say" is counted as an answer.

### Age

Respondents skew older: most are 55 or over.

```{r}
#| label: age
demographics |>
  filter(!is.na(age)) |>
  count(age) |>
  ggplot(aes(n, fct_rev(age))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.3, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(title = "Age", x = "Respondents", y = NULL)
```

### Race and ethnicity

People could choose more than one answer, so shares add to more than 100
percent.

```{r}
#| label: race
demographics |>
  selection_rates("race") |>
  plot_rates("Race and ethnicity", "Share of respondents choosing each; more than one allowed")
```

### Gender

```{r}
#| label: gender
demographics |>
  selection_rates("gender") |>
  transmute(gender = label, respondents = n_selected, share = percent(rate, accuracy = 1)) |>
  arrange(desc(respondents)) |>
  knitr::kable()
```

### Household income

```{r}
#| label: income
share_table(demographics, hh_income)
```

### Household size

```{r}
#| label: hh-size
share_table(demographics, hh_size)
```

### Disability or chronic health condition

```{r}
#| label: disability
share_table(demographics, disability)
```

### Political views

```{r}
#| label: political
share_table(demographics, political)
```

## Connection to care

The first question asks about a person's connection to care and routes them
to one of five pathways. This decides which questions each person sees, so it
shapes everything below. Current caregivers and care recipients are by far
the largest group.

```{r}
#| label: pathways
bind_rows(survey = survey, canvassing = canvassing, .id = "tool") |>
  filter(!is.na(pathway)) |>
  count(tool, pathway) |>
  mutate(share = percent(n / sum(n), accuracy = 1), .by = tool) |>
  pivot_wider(names_from = tool, values_from = c(n, share)) |>
  select(pathway, survey_n = n_survey, survey_share = share_survey,
         canvassing_n = n_canvassing, canvassing_share = share_canvassing) |>
  knitr::kable()
```

The answers behind the pathways. People could choose more than one.

```{r}
#| label: q1
#| fig-height: 5
survey |>
  selection_rates("q1") |>
  plot_rates(
    "Connection to care (survey)",
    str_glue("Share of {nrow(filter(survey, !is.na(q1_child_now)))} respondents who answered; more than one allowed")
  )
```

## What people are saying

For each question, the chart shows the share of people *who were shown the
question* choosing each answer, with the number of people in parentheses.
Because the pathways see different questions, the group behind each chart is
different; the subtitle says who.

### Challenges for current caregivers and care recipients

```{r}
#| label: q2a
#| fig-height: 5
q2a <- selection_rates(survey, "q2a")
plot_rates(
  q2a,
  "What challenges have you experienced with care?",
  str_glue("Survey, current pathway, {unique(q2a$n_shown)} respondents")
)
```

### Supports that would make a difference for current caregivers and recipients

```{r}
#| label: q3a
#| fig-height: 5
q3a <- selection_rates(survey, "q3a")
plot_rates(
  q3a,
  "What care support would make a difference for you?",
  str_glue("Survey, current pathway, {unique(q3a$n_shown)} respondents")
)
```

### Who helps most with care

```{r}
#| label: q4a
#| fig-height: 5
q4a <- selection_rates(survey, "q4a")
plot_rates(
  q4a,
  "What or who helps you most with care? (top 3)",
  str_glue("Survey, current pathway, {unique(q4a$n_shown)} respondents")
)
```

### What ideal care would look like

Everyone sees this question regardless of pathway, so it has the widest base.
It asks people to choose up to three answers, but the form does not enforce
that, and about a third of respondents chose more — some chose all eleven.
The table shows shares two ways: among everyone, and among only the people
who kept to three or fewer. The ranking is similar either way, but the
levels differ, and it is worth deciding which version to report.

```{r}
#| label: q6
#| fig-height: 5.5
q6_all <- selection_rates(survey, "q6")

q6_n_chosen <- survey |>
  filter(!is.na(q6_affordable)) |>
  mutate(n_chosen = rowSums(across(starts_with("q6_") & where(is.logical) & !q6_other))) |>
  select(respondent_id, n_chosen)

q6_within <- survey |>
  semi_join(filter(q6_n_chosen, n_chosen <= 3), by = "respondent_id") |>
  selection_rates("q6")

plot_rates(
  q6_all,
  "What would ideal care support and services look like for you?",
  str_glue("Survey, all pathways, {unique(q6_all$n_shown)} respondents")
)
```

```{r}
#| label: q6-table
q6_all |>
  select(label, everyone = rate) |>
  left_join(select(q6_within, label, three_or_fewer = rate), by = "label") |>
  arrange(desc(everyone)) |>
  mutate(across(c(everyone, three_or_fewer), \(x) percent(x, accuracy = 1))) |>
  rename(
    Answer = label,
    "Everyone ({unique(q6_all$n_shown)})" := everyone,
    "Chose 3 or fewer ({unique(q6_within$n_shown)})" := three_or_fewer
  ) |>
  knitr::kable()
```

### Ideal care by respondent group

The same question, split by age and by disability status. Each column is one
group; the number in the heading is how many people in that group answered.
Groups with fewer than 30 respondents are left out. With more responses these
splits can extend to every question and every demographic.

```{r}
#| label: q6-by-age
rates_by(survey, "q6", age) |>
  knitr::kable()
```

```{r}
#| label: q6-by-disability
rates_by(survey, "q6", disability) |>
  knitr::kable()
```

### In their own words

Every choose-all-that-apply question has an "Other" box. Text written there
is kept word for word and will be coded alongside the open-ended questions.
Number of respondents who wrote something, by question:

```{r}
#| label: other-text
survey |>
  select(ends_with("_other_text")) |>
  summarise(across(everything(), \(x) sum(!is.na(x)))) |>
  pivot_longer(everything(), names_to = "question", values_to = "respondents") |>
  mutate(question = str_remove(question, "_other_text")) |>
  filter(respondents > 0) |>
  arrange(desc(respondents)) |>
  knitr::kable()
```

## Canvassing

Canvassing is a doorstep conversation, so it is a funnel: not everyone
approached answers, not everyone who answers has a connection to care, and
not everyone with a connection agrees to the full questions. Each stage is a
different base for any analysis.

```{r}
#| label: funnel
funnel <- tibble(
  stage = c(
    "Approached",
    "Answered and engaged",
    "Reports a connection to care",
    "Agreed to full questions"
  ),
  n = c(
    nrow(canvassing),
    sum(canvassing$responded == "Yes - continue with questions", na.rm = TRUE),
    sum(canvassing$care_connection == "Yes", na.rm = TRUE),
    sum(!is.na(canvassing$pathway))
  )
) |>
  mutate(stage = fct_rev(fct_inorder(stage)))

ggplot(funnel, aes(n, stage)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.5, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(title = "Canvassing doorstep funnel", x = "Contacts", y = NULL)
```

At the door, the canvasser records challenges from a short list before the
full questions begin. This covers everyone who engaged, not just those who
went on to the full questions.

```{r}
#| label: doorstep
doorstep <- selection_rates(canvassing, "doorstep")
plot_rates(
  doorstep,
  "Challenges mentioned at the door",
  str_glue("Canvassing, {unique(doorstep$n_shown)} engaged contacts")
)
```

The full question set was answered by too few canvassing contacts to chart
yet (`r sum(!is.na(canvassing$q1_child_now))` so far). Those answers are in
the clean file and will be added here as canvassing grows.

## Interviews

Three interview transcripts have been delivered: one group session and two
individual interviews. All with a ZIP code are in Illinois. Interview
content is being coded against the project codebook and reported separately
in the coding dashboard.

```{r}
#| label: interviews
interviews |>
  left_join(zip_state, by = "zip") |>
  left_join(states, by = "state") |>
  select(transcript, format, zip, state = state_name) |>
  knitr::kable()
```

## Not in this chartbook yet

- **Rural detail.** Rural respondents can be split further (small towns,
  remote areas, commuting patterns) once there are enough of them.
- **Open-ended answers.** The survey's and canvassing's open-text questions
  are coded separately once the codebook is updated (issue #2).
- **Later channels.** The email opt-in, Spanish, and film-screening survey
  links and the second canvassing event have not been delivered yet.
- **Comparison to the population.** Demographic shares above are not yet
  compared to a national benchmark.
