---
title: "Cleaning the survey and canvassing data"
author: "Manuel Alcalá Kovalski, Urban Institute"
date: today
execute:
  warning: true
  message: false
---

*Program that turns the raw survey and canvassing exports from Caring Across
Generations into clean, analysis-ready files.*

The raw exports arrive as Excel files in a Box folder. `scripts/00_ingest-raw.R`
copies that folder to `data/raw/Raw data backups/` exactly as it is on Box.
This program reads that local copy, cleans it, and writes three files to
`data/processed/`:

- `survey_clean.rds` / `.csv` — one row per survey respondent
- `canvassing_clean.rds` / `.csv` — one row per canvassing contact
- `data-dictionary.csv` — every column in those files, with the question and
  answer text it came from

## The steps at a glance

1. **Find the files.** Pick the survey and canvassing exports out of the Box
   mirror by their file names; set aside the test delivery, the open-text
   exports, and the pre-launch poll. Every file must be claimed by a rule.
2. **Rename the columns** from full question text to short names (`q1`,
   `q2a`, `age`, ...), checking that nothing is missing or unexpected.
3. **Check for personal information.** Address columns must be empty.
4. **Split the choose-all-that-apply questions** into one TRUE/FALSE column
   per answer, keeping "Other" text word for word.
5. **Put single-answer questions in order** (income, age, and so on).
6. **Work out each person's pathway** through the survey from their first
   answer, and **check** that they answered the questions for that pathway.
7. **Tidy the rest**: IDs, timestamps, ZIP codes. Confirm no ID repeats.
8. **Do the same for canvassing**, reusing the survey's answer lists.
9. **Write the clean files** and a data dictionary describing every column.

::: {.callout-warning title="Warnings in this document are expected"}
Every step checks its own work. If a delivery looks different from what we
expect — a renamed column, an answer we have not seen before, a file no rule
claims — the program stops and says why, rather than guessing. The fix is
always to update the rule that failed, never to loosen the check.

Some checks warn instead of stopping; those warnings appear below marked with
`!`. The ones documented in `SURVEY.md` (the paid-provider routing exception,
ZIP+4 entries, one canvassing funnel violation) are expected. A *new* warning
is the signal to investigate.
:::

::: {.callout-note title="The live form does not match the questionnaire PDF"}
The wording in the live form differs from the official question documents in
many places, and the export format itself has changed between deliveries.
Answer text in this program comes from what we actually observe in the
exports, not from the PDF. `SURVEY.md` records what we have learned about how
the instrument really behaves.
:::

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(readxl)
library(here)
library(gt)

source(here("scripts/00_utils.R"))

# consistent look for the small tables in this document
tbl <- function(data, title = NULL) {
  out <- gt(data) |>
    sub_missing(missing_text = "—") |>
    tab_options(table.font.size = px(14), table.align = "left")
  if (!is.null(title)) out <- tab_header(out, title = title) |> opt_align_table_header("left")
  out
}
```

## Find the files to clean

The file names in the Box folder already tell us what each file is: which
tool (`surv` or `canv`), the date range, the distribution channel (`f1`, `f2`,
...), whether it holds closed-ended or open-text answers, and whether the
responses are complete or partial. We select files by those name patterns and
never rename anything by hand.

Not every file is cleaned here. The `f0` files are the pre-launch test and are
left out. The open-text exports and the pre-launch poll go to the
text-analysis pipeline instead. Every file in the folder has to fall into one
of three groups — cleaned here, deliberately set aside, or an error — so a new
kind of delivery can never slip through unnoticed.

```{r}
#| label: read
raw_dir <- here("data/raw/Raw data backups")

mirror_files <- list.files(raw_dir) |>
  str_subset("~\\$", negate = TRUE) # Excel lock files, present while a workbook is open

# deliveries from 2026-08 on carry a "nopii_" prefix (PII-skimmed upstream);
# f0 = pre-launch test deliveries, excluded per the data log
survey_files <- mirror_files |>
  str_subset("^(nopii_)?surv_export_.*_closed.*\\.xlsx$") |>
  str_subset(regex("_f0_", ignore_case = TRUE), negate = TRUE)
canvassing_files <- mirror_files |>
  str_subset("^(nopii_)?canv_export_.*_closed.*\\.xlsx$") |>
  str_subset(regex("_f0_", ignore_case = TRUE), negate = TRUE)

set_aside <- c(
  "_[fF]0_",     # pre-launch test deliveries (both tools)
  "opentext",    # open-text exports — text-analysis pipeline, not read here
  "_open\\.",    # ... newer deliveries shorten the suffix to "_open"
  "everyaction", # pre-launch One Question Poll — text-analysis pipeline
  "grp-sf",      # interview holding document
  "DATA LOG"     # the team's delivery manifest
)
unaccounted <- mirror_files |>
  setdiff(c(survey_files, canvassing_files)) |>
  str_subset(str_flatten(set_aside, "|"), negate = TRUE)
if (length(unaccounted) > 0) {
  cli::cli_abort(c(
    "{length(unaccounted)} file{?s} in the mirror not claimed by a cleaning section or known set-aside:",
    set_names(unaccounted, rep("x", length(unaccounted)))
  ))
}

survey_files <- file.path(raw_dir, survey_files)
canvassing_files <- file.path(raw_dir, canvassing_files)
```

These are the survey files cleaned in this run. Each one should appear in the
team's data log.

```{r}
#| label: read-list
#| echo: false
basename(survey_files)
```

## Survey

### Give the columns short names

In the export, each column is named with the full question text. We map those
to short names that follow the questionnaire's numbering: `q1` is the
connection-to-care question, `q2a`/`q3a`/`q4a` are the current-caregiver
questions, and so on through `q6`. Demographic columns get plain names such as
`age` and `hh_income`.

The mapping is checked in both directions. A column in the export that we have
no name for stops the program, and so does a column we expect but cannot find.
A few columns are allowed to be missing because they come and go across
deliveries (the empty address fields in the test delivery, and `q4c`, which
the test delivery lacked).

The open-text questions (q5 and q7) are not in these files. They arrive in a
separate export and are handled by the text-analysis pipeline.

```{r}
#| label: rename
col_map <- c(
  submitted_at  = "Time",
  q1            = "We would like to know more about your connection to care, if any. Choose all that apply to you.",
  q2a           = "What challenges, if any, have you experienced with providing, finding, or receiving care? (Choose all that apply)",
  q3a           = "What care support would make a difference for you or people around you? (Choose all that apply)",
  q4a           = "What or who helps you most with care? (Choose your top 3)",
  q2b           = "When you think about the care you, a family member, or friend might need 5 years from now, what challenges, if any, concern you most? (Choose all that apply)",
  q3b           = "What types of care support and services do you think will matter most for you or people you know? (Choose all that apply)",
  q2c           = "When you provided or received care what challenges, if any, did you experience? (Choose all that apply)",
  q3c           = "What care support would have made a difference for you or people around you? (Choose all that apply)",
  q2d           = "What care challenges, if any, do the people you know experience? (Choose all that apply.)",
  q3d           = "What care support would make a difference for the people you know? (Choose all that apply)",
  q4d           = "What or who helps most with their care? (Choose your top 3)",
  q4c           = "What or who helped you most with care? (Choose your top 3)",
  q6            = "What would ideal care support and services look like for you? (Choose up to three)",
  address       = "Zipcode (Address)",
  address2      = "Zipcode (Address2)",
  city          = "Zipcode (City)",
  state         = "Zipcode (State)",
  zip           = "Zipcode (Zip)",
  hh_size       = "How many people live in your household? (Include yourself and anyone you share income and living expenses with, including kids)",
  hh_income     = "What is your household's approximate annual income? (Add up income from everyone living with you, related or not.)",
  age           = "What is your age?",
  gender        = "What is your gender? (Choose all that apply)",
  race          = "What is your race or ethnicity? (Choose all that apply)",
  disability    = "I identify as a person with a disability or chronic health condition.",
  political     = "How would you describe your political views?",
  respondent_id = "Unique ID"
)

form_varying <- c("address", "address2", "city", "state", "q4c")

# exports from Aug 2026 on label the timestamp column "Date submitted"
# instead of "Time"; same Excel serial values underneath
timestamp_aliases <- c("Time" = "Date submitted")

# the same export writes en-/em-dashes and curly apostrophes ("$25,000–$49,999",
# "Affordable — without ...", "I don’t") where every other export writes plain
# hyphens and straight quotes
normalize_punctuation <- \(x) str_replace_all(x, c("—" = "-", "–" = "-", "’" = "'", "‘" = "'"))

read_delivery <- function(path) {
  read_excel(path, col_types = "text") |>
    rename(any_of(timestamp_aliases)) |>
    mutate(across(everything(), \(x) na_if(str_trim(normalize_punctuation(x)), ""))) |> # whitespace-only cells -> NA
    rename_validated(col_map, optional = form_varying) |>
    mutate(delivery = basename(path))
}

survey <- survey_files |>
  map(read_delivery) |>
  bind_rows() |>
  mutate(
    # filenames are inconsistent about case ("F1")
    form_version    = str_to_lower(str_extract(delivery, regex("f\\d+", ignore_case = TRUE))),
    response_status = str_extract(delivery, "complete|partial")
  )

# the filename f0-exclusion and the parsed form_version must agree
if (any(survey$form_version == "f0")) {
  cli::cli_abort("f0 test rows slipped past the filename exclusion")
}
```

Count the rows that came from each file. The `form_version` column records
the distribution channel (the May launch link, the July Virtual Day of Action
link, postcards, and so on — see `SURVEY.md`); every channel uses the same
questionnaire. `response_status` separates people who finished the survey from
those who stopped partway through. Both are kept in the clean file so the
analysis can decide how to treat them.

```{r}
#| label: read-count
#| echo: false
count(survey, delivery, form_version, response_status) |>
  tbl("Rows read from each delivery")
```

This is what the raw export looks like at this point: one row per person,
the timestamp as an Excel serial number, and every choose-all-that-apply
answer packed into a single cell. The two rows below come from two different
deliveries and show the two formats the export has used — vertical bars, and
(in the July–August complete file) commas.

```{r}
#| label: raw-sample
#| echo: false
sample_rows <- c(
  which(str_detect(survey$q2a, fixed("|")))[1],
  which(str_detect(survey$delivery, "20260710-20260812_f1_complete") & !is.na(survey$q2a))[1]
)

survey[sample_rows, c("delivery", "submitted_at", "q1", "q2a", "hh_income")] |>
  mutate(across(c(q1, q2a), \(x) str_trunc(x, 110))) |>
  tbl("Two raw rows, selected columns")
```

### Check for personal information

::: {.callout-important title="No personal information"}
The data shared with Urban must contain no personal information. ZIP code is
collected on purpose; street address is not. The export still carries address
columns, which should always be empty. If they ever are not, the program stops
here so the file can be raised with the team before anything else happens.
:::

```{r}
#| label: pii-guard
pii_cols <- intersect(c("address", "address2"), names(survey))
if (length(pii_cols) > 0) {
  if (any(!is.na(as.matrix(survey[pii_cols])))) {
    stop("Address fields are not empty — possible PII in the raw file. Do not proceed; flag to the team.")
  }
  survey <- select(survey, -all_of(pii_cols))
}
```

### Split the choose-all-that-apply questions into one column per answer

Most survey questions let people choose more than one answer. The export
stores all of a person's choices in a single cell, usually separated by
vertical bars, for example `"Care is unaffordable|Other: my own words"`. One
delivery (the July–August complete responses) separates them with commas
instead, which is trickier because the answers themselves contain commas; for
those cells we match the known answers one at a time from the start of the
cell. Either way, the result is the same: each question is split into one
TRUE/FALSE column per answer option.
For the challenges question `q2a`, that means columns such as
`q2a_unaffordable` and `q2a_time_flex`.

::: {.callout-tip title="How to read the TRUE/FALSE columns"}
Each answer column holds one of three values:

- `TRUE` — the person chose this answer
- `FALSE` — the person saw the question and did not choose this answer
- `NA` — the person never saw the question (the survey routed them past it) or
  left it blank

The share of people choosing an answer is simply the share of `TRUE` among
the non-missing values.
:::

"Other" works a little differently. It is a free-text box, and the export
records it three ways: `"Other: some text"`, a bare `"Other"` when the box
was ticked but left empty, or occasionally the text alone. Anything that is
not one of the known answers sets `{question}_other` to `TRUE` and is kept
word for word in `{question}_other_text`. Those words are respondents'
own voices and feed the text analysis.

Two safeguards sit inside this step. If the same "other" text appears from
three or more people, the program warns — repeated text usually means a new
answer option was added to the form and needs to be added to our list. And
when the form's wording for an answer changes (as it did on August 19, 2026,
for four options), both the old and the new wording are listed for that
answer, so responses from before and after the change land in the same
column.

A small demonstration on a made-up question with answers `a` through `f`
shows how the pieces fit: `"c | e"` and the comma form `"c, e"` both get
`TRUE` for c and e and `FALSE` for the rest; an "Other" with text sets the
other flag and keeps the text; a blank row is `NA` throughout; and answer `a`
accepts two wordings.

```{r}
#| label: encode-demo
tibble(q_demo = c("c | e", "c, e", "a-reworded|Other: my own answer", "a", NA)) |>
  encode_multiselect(
    "q_demo",
    list(a = c("a", "a-reworded"), b = "b", c = "c", d = "d", e = "e", f = "f")
  )
```

The answer options for every question are listed below, in the wording the
export uses. The short name on the left becomes the column suffix. Where the
live form does not yet match the official documents, the observed wording
wins.

```{r}
#| label: dictionaries
multiselect_options <- list(
  q1 = c(
    child_now      = "I care for a child or children now",
    aging_now      = "I care for an aging person now",
    disability_now = "I care for someone with a disability or chronic health condition now",
    paid_provider  = "I am a paid care provider. I offer support and services for kids, aging adults, and/or people with disabilities",
    need_care_now  = "I need care but am not receiving it",
    receives_care  = "I receive or have received care for a disability or chronic condition",
    cared_past     = "I have cared for an aging adult, child, or someone with a disability or chronic health condition in the past",
    expect_future  = "I am not providing or receiving care right now, but I expect to in the next 5 years",
    observer       = "None of these apply to me right now, but I know someone who receives or needs care",
    none           = "None of these apply to me"
  ),
  q2a = list(
    unaffordable = "Care is unaffordable",
    finding      = c(
      "It is hard to find care that's available, nearby, or right for my family or me",
      # Aug 19 2026 revision — seeded from the questionnaire PDF, not yet observed in an export
      "It's hard to find care that's available, nearby, or right for me or my family"
    ),
    time_flex    = "I don't have enough time or flexibility",
    help         = "I don't have enough help or support",
    figure_out   = "It is hard to figure out what help exists or how to get it",
    navigate     = "It is hard to navigate existing care programs",
    no_say       = "I don't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3a = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    everyday_support   = "Everyday support like meals and transportation",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4a = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, child care, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid for using federal, state, or county dollars",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I do not receive help"
  ),
  q2b = c(
    unaffordable = "Care that's unaffordable",
    finding      = "Difficulty with finding care that's available, nearby, or right for me or my family",
    time_flex    = "Not having time or flexibility",
    help         = "Not enough help or support with care",
    figure_out   = "Difficulty knowing what help exists or how to get it",
    navigate     = "Difficulty navigating existing care programs",
    no_say       = "Not having a say in how care works for my family or me",
    job          = "Care could affect my job or career"
  ),
  q3b = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q2c = c(
    unaffordable = "Care was unaffordable",
    finding      = "It was hard to find care that was available, nearby, or right for my family and me",
    time_flex    = "I didn't have enough time or flexibility",
    help         = "I didn't have enough help or support",
    figure_out   = "It was hard to figure out what help existed or how to get it",
    navigate     = "It was hard to navigate existing care programs",
    no_say       = "I didn't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3c = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4c = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  q2d = c(
    unaffordable = "Unaffordable care",
    finding      = "It was hard to find care that's available, nearby, or right for them",
    time_flex    = "They don't have enough time or flexibility",
    help         = "They are doing this without enough help or support",
    figure_out   = "It is hard for them to figure out what help exists or how to get it",
    navigate     = "It is hard for them to navigate existing care programs",
    no_say       = "They don't have a say in how care works for them",
    job          = "Care has affected their jobs or careers"
  ),
  q3d = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4d = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    # in the instrument (per the Aug 19 2026 PDF, wording as in q4c) but unobserved so far
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  # Aug 19 2026 variants are seeded from the questionnaire PDF with the
  # formatting pattern of observed exports (plain hyphens, not the PDF's
  # em-dashes) — confirm against the first export that carries them
  q6 = list(
    there_when_needed = "There when you need it - at every stage of life, no matter your job or family situation",
    easy_to_use       = c(
      "Easy to find and use - without complicated steps or barriers",
      "Easy to find and use - without confusing steps or barriers"          # Aug 19 2026
    ),
    affordable        = "Affordable - without causing personal financial strain",
    in_community      = "Available in your community - not far away or out of reach",
    flexible          = "Flexible - fitting real schedules, needs, and changing circumstances",
    reliable          = "Reliable - something you can count on over time",
    clear             = "Clear and transparent - you know what's available and how to get it",
    good_jobs         = "Good jobs for care providers - with fair pay and stability",
    healthy_dev       = c(
      "Support healthy development, well-being, and independence",
      "Supports healthy development, well-being, and independence"          # Aug 19 2026
    ),
    lived_experience  = c(
      "Shaped by people with experience giving, receiving, or needing care",
      "Shaped by people with lived care experience"                         # Aug 19 2026
    ),
    secure            = "Keep people who rely on care feeling secure, healthy, and thriving"
  ),
  gender = c(
    woman             = "Woman",
    man               = "Man",
    nonbinary         = "Non-binary",
    transgender       = "Transgender",
    prefer_not_to_say = "Prefer not to say"
  ),
  race = c(
    aian              = "American Indian or Alaska Native",
    asian             = "Asian or Asian American",
    black             = "Black or African American",
    hispanic          = "Hispanic, Latino, or Latina",
    mena              = "Middle Eastern or North African",
    nhpi              = "Native Hawaiian or Pacific Islander",
    white             = "White",
    multiracial       = "Multiracial",
    prefer_not_to_say = "Prefer not to say"
  )
)

survey <- reduce(
  names(multiselect_options),
  \(data, col) encode_multiselect(data, col, multiselect_options[[col]]),
  .init = survey
)
```

The same two rows as before, after this step, for the challenges question
`q2a`. The packed cell has become ten TRUE/FALSE columns plus the "Other"
text.

```{r}
#| label: encoded-sample
#| echo: false
survey[sample_rows, ] |>
  select(starts_with("q2a_")) |>
  pivot_longer(everything(), names_to = "column", values_to = "value", values_transform = as.character) |>
  mutate(row = rep(c("row 1", "row 2"), each = n() / 2), .before = 1) |>
  pivot_wider(names_from = row, values_from = value) |>
  tbl("The q2a columns for the two sample rows")
```

::: {.callout-note title="q6 asks for up to three answers, but the form does not enforce it"}
The ideal-care question `q6` asks people to choose up to three answers. The
live form does not enforce that limit, and many people choose more. We keep
every selection as given; the chartbook reports `q6` both ways. The count
below shows how many answers people actually chose.
:::

```{r}
#| label: q6-limit
#| echo: false
survey |>
  filter(!is.na(q6_affordable)) |>
  mutate(n_chosen = rowSums(across(starts_with("q6_") & where(is.logical) & !q6_other))) |>
  count(n_chosen, name = "respondents") |>
  mutate(share = scales::percent(respondents / sum(respondents), accuracy = 1)) |>
  tbl("Number of q6 answers chosen")
```

### Put single-answer questions in order

The demographic questions with one answer each become ordered categories
(household size from smallest to largest, income from lowest to highest, and
so on) so that tables and charts sort sensibly. Levels the export has not
shown us yet are filled in from the questionnaire, following the format of
the values we have seen. An answer outside the expected set stops the program.

```{r}
#| label: factors
survey <- survey |>
  mutate(
    # the export writes household size as "2.0", "3.0", ...
    hh_size = to_factor(str_remove(hh_size, "\\.0$"), c(
      "1, just myself", "2", "3", "4", "5", "6", "7 or more",
      "Prefer not to say"
    )),
    hh_income = to_factor(hh_income, c(
      "Under $25,000", "$25,000-$49,999", "$50,000-$74,999",
      "$75,000-$99,999", "$100,000-$149,999", "$150,000-$174,999",
      "$175,000-$199,999", "$200,000-$249,999", "$250,000-$299,999",
      "$300,000 or more", "Prefer not to say"
    )),
    age = to_factor(age, c(
      "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older",
      "Prefer not to say"
    )),
    disability = to_factor(disability, c(
      "Yes", "No", "Not sure", "Prefer not to say"
    )),
    political = to_factor(political, c(
      "Very liberal", "Somewhat liberal", "Moderate/Middle of the road",
      "Somewhat conservative", "Very conservative", "Don't know",
      "Prefer not to say"
    ))
  )
```

### Work out each person's pathway

The first question asks about a person's connection to care, and the answer
decides which questions they see next. There are five pathways: people giving
or receiving care now (`current`), people who cared for someone in the past
(`past`), people who expect to in the future (`future`), people who know
someone who receives care (`observer`), and people with no connection
(`none`). A person who ticks answers from more than one group is routed to
the first that applies, in that order. We checked this against the data — for
example, a person eligible for both `past` and `future` was shown only the
`past` questions.

The survey's routing, as it behaves in the data:

```{mermaid}
%%| label: fig-routing
%%| fig-cap: "How respondents move through the survey. The first pathway that applies wins."
flowchart TD
  q1["q1 · Connection to care<br/>(choose all that apply)"]
  q1 -->|"caring for someone now, paid provider,<br/>or needs / receives care"| A
  q1 -->|"cared for someone in the past"| C
  q1 -->|"expects to give or need care"| B
  q1 -->|"knows someone who receives care"| D
  q1 -->|"none of these"| E

  A["current<br/>q2a challenges · q3a supports · q4a who helps"]
  C["past<br/>q2c · q3c · q4c"]
  B["future<br/>q2b · q3b"]
  D["observer<br/>q2d · q3d · q4d"]
  E["none"]

  A --> q6
  C --> q6
  B --> q6
  D --> q6
  E --> q6
  q6["q6 · Ideal care (everyone)"] --> demo["Demographics<br/>ZIP · household · income · age · gender<br/>race · disability · political views"]
```

The open-ended questions (q5 on three of the pathways, q7 for everyone)
sit between q4 and q6 but arrive in a separate export, so they are not shown.

```{r}
#| label: pathway
survey <- derive_pathway(survey)
```

Count respondents on each pathway. `NA` here means the person stopped before
answering the first question.

```{r}
#| label: pathway-count
#| echo: false
count(survey, pathway, name = "respondents") |>
  tbl("Respondents by pathway")
```

### Check that people answered the questions for their pathway

Each pathway question should be answered by exactly the people on that
pathway. For complete responses we check both ways: nobody on the pathway
should have skipped the question, and nobody off it should have answered.
Partial responses may stop early, so they are only flagged if they answered a
question outside their pathway. Mismatches produce a warning rather than
stopping the program, because they point to a routing or export question for
the team rather than a cleaning error. The known exception is documented in
`SURVEY.md`: a few paid care providers with past experience were shown two
sets of questions.

```{r}
#| label: skip-checks
skip_rules <- tribble(
  ~question, ~indicator,             ~expected_pathway,
  "q2a",     "q2a_unaffordable",     "current",
  "q3a",     "q3a_backup",           "current",
  "q4a",     "q4a_family_friends",   "current",
  "q2b",     "q2b_job",              "future",
  "q3b",     "q3b_financial",        "future",
  "q2c",     "q2c_job",              "past",
  "q3c",     "q3c_backup",           "past",
  "q4c",     "q4c_family_friends",   "past",
  "q2d",     "q2d_unaffordable",     "observer",
  "q3d",     "q3d_backup",           "observer",
  "q4d",     "q4d_family_friends",   "observer"
)

# q6 is shown to every pathway, so it has no row in skip_rules — instead a
# complete response that skipped it entirely signals routing/export drift
q6_skipped <- survey$response_status == "complete" & is.na(survey$q6_affordable)
if (any(q6_skipped)) {
  cli::cli_warn("q6: {sum(q6_skipped)} complete response{?s} with q6 entirely missing")
}

pwalk(skip_rules, \(question, indicator, expected_pathway) {
  answered <- !is.na(survey[[indicator]])
  eligible <- survey$pathway %in% expected_pathway
  mismatch <- if_else(
    survey$response_status == "complete",
    xor(answered, eligible), # completes must answer exactly their pathway
    answered & !eligible     # partials may stop early, but not cross paths
  )
  n_bad <- sum(mismatch, na.rm = TRUE)
  if (n_bad > 0) {
    cli::cli_warn(
      "{question}: {n_bad} respondent{?s} answered or skipped against skip logic"
    )
  }
})
```

### Tidy the remaining columns

Three small fixes. Respondent IDs sometimes export in scientific notation, so
they are rewritten as plain digits (this also lets us confirm no ID appears
twice). The submission time arrives as an Excel serial number and is converted
to a date-time. ZIP codes pick up a stray `.0` from the export and lose their
leading zeros, so both are repaired; any ZIP that still is not five digits is
flagged.

```{r}
#| label: other-cols
survey <- survey |>
  mutate(
    respondent_id = normalize_id(respondent_id),
    # tz assumed UTC — the platform's actual export timezone is unconfirmed
    submitted_at = excel_datetime(submitted_at),
    zip = str_pad(str_remove(zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

# after normalization so ID collisions across export formats would surface
dupes <- survey$respondent_id[duplicated(survey$respondent_id)]
if (length(dupes) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes)}}")
}

bad_zip <- survey |> filter(!is.na(zip), !str_detect(zip, "^\\d{5}$"))
if (nrow(bad_zip) > 0) {
  cli::cli_warn("{nrow(bad_zip)} row{?s} with a ZIP that is not five digits (ZIP+4 entries, kept as typed)")
}
```

Missing ZIPs are expected among partial responses: ZIP is the first
demographic question, and many partials stop before reaching it.

```{r}
#| label: zip-count
#| echo: false
count(survey, response_status, zip_missing = is.na(zip), name = "respondents") |>
  tbl("ZIP code present, by response status")
```

```{r}
#| label: write
#| include: false
write_rds(survey, here("data/processed/survey_clean.rds"))
write_csv(survey, here("data/processed/survey_clean.csv"))
```

The clean survey file has one row per respondent and
`r ncol(survey)` columns. Most are the TRUE/FALSE answer columns; the rest
identify the respondent and delivery or hold the single-answer demographics.

```{r}
#| label: survey-structure
#| echo: false
tibble(column = names(survey), type = map_chr(survey, \(x) class(x)[1])) |>
  mutate(
    group = case_when(
      str_detect(column, "^(q\\d[a-d]?|gender|race)_") ~ str_extract(column, "^[^_]+"),
      TRUE ~ "respondent / delivery"
    ),
    type = case_match(type, "logical" ~ "TRUE/FALSE", "factor" ~ "ordered category",
                      "character" ~ "text", "POSIXct" ~ "date-time", .default = type)
  ) |>
  summarise(columns = n(), types = str_flatten(unique(type), ", "),
            examples = str_flatten(head(column, 3), ", "), .by = group) |>
  tbl("Columns in the clean survey file")
```

## Canvassing

The canvassing form is used at people's doors. It starts with a few questions
for the canvasser (where they are, how the conversation is happening, whether
the person answered), a short screener, and a quick-pick list of care
challenges. People who agree to continue then get the same question set as
the survey, word for word, so the survey's column names and answer lists are
reused. There are no demographic questions and no respondent ZIP; the only
geography is the canvasser's ZIP.

As with the survey, the `f0` test delivery is left out, and complete and
partial responses are combined and labelled.

```{r}
#| label: canvassing-read
#| echo: false
basename(canvassing_files)
```

```{r}
#| label: canvassing-rename
battery <- col_map[c(
  "q1", "q2a", "q3a", "q4a", "q2b", "q3b", "q2c", "q3c",
  "q2d", "q3d", "q4d", "q6"
)]

col_map_canv <- c(
  submitted_at        = "Time",
  canvasser_zip       = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Zip)",
  canvass_mode        = "For Canvasser: How are you conducting this conversation?",
  responded           = "For Canvasser: Did the person respond and agree to answer questions?  If no, share the information card or the careconversations.com website and submit.",
  care_connection     = "Have you provided, received care yourself, expected to in the future or do you know someone who is? This could be giving or receiving care for an aging family member, a friend, a child, or someone with a disability no matter their age. It also includes you.",
  doorstep             = "What challenges, if any, have you or others faced — or do you expect to face — when it comes to giving, finding, or getting care? (For canvasser: Check all that apply. If selecting 'No Challenges,' leave everything else unchecked.)",
  continue_full       = "We're talking with people across the country about their experiences with care — what's working and what families need.  Can you take 5-minutes to share more?",
  battery,
  q4c                 = "What or who helped you most with care? (Choose your top 3)",
  respondent_id       = "Unique ID"
)

canvassing <- canvassing_files |>
  map(\(path) {
    read_excel(path, col_types = "text") |>
      rename(any_of(timestamp_aliases)) |>
      mutate(across(everything(), \(x) na_if(str_trim(x), ""))) |>
      rename_validated(col_map_canv) |>
      mutate(delivery = basename(path))
  }) |>
  bind_rows() |>
  mutate(
    form_version    = str_extract(delivery, "f\\d+"),
    response_status = str_extract(delivery, "complete|partial")
  )

if (any(canvassing$form_version == "f0")) {
  cli::cli_abort("f0 test rows slipped past the filename exclusion")
}
```

```{r}
#| label: canvassing-count
#| echo: false
count(canvassing, delivery, form_version, response_status) |>
  tbl("Rows read from each canvassing delivery")
```

The doorstep challenge list uses short labels of its own. Everything else
reuses the survey answer lists.

```{r}
#| label: canvassing-encode
doorstep_options <- c(
  cost                = "Cost",
  time                = "Time",
  location            = "Location",
  right_fit           = "Right fit",
  lack_support        = "Lack of support",
  no_control          = "No control",
  worker_availability = "Worker availability",
  barriers_access     = "Barriers to access",
  no_challenges       = "No challenges",
  # added to the form between f0 and f1 (surfaced by the repeat-drift alarm)
  availability        = "Availability",
  figure_out          = "Hard to figure out"
)

canv_multiselect <- c(
  list(doorstep = doorstep_options),
  multiselect_options[c(
    "q1", "q2a", "q3a", "q4a", "q2b", "q3b",
    "q2c", "q3c", "q4c", "q2d", "q3d", "q4d", "q6"
  )]
)

canvassing <- reduce(
  names(canv_multiselect),
  \(data, col) encode_multiselect(data, col, canv_multiselect[[col]]),
  .init = canvassing
) |>
  derive_pathway()
```

The same tidying as the survey, plus one check specific to canvassing: the
full question set should be answered by exactly the people who agreed to
continue. As before, partials may stop early but should not answer questions
without agreeing.

```{r}
#| label: canvassing-clean
canvassing <- canvassing |>
  mutate(
    respondent_id   = normalize_id(respondent_id),
    canvass_mode    = to_factor(canvass_mode, c("Door-to-door", "Phone", "Other In Person")),
    responded       = to_factor(responded, c("Yes - continue with questions", "No - not available")),
    care_connection = to_factor(care_connection, c("Yes", "No")),
    continue_full   = to_factor(continue_full, c("Yes - continuing", "No - not continuing")),
    submitted_at    = excel_datetime(submitted_at),
    # same float artifact as the survey ZIPs
    canvasser_zip   = str_pad(str_remove(canvasser_zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

dupes_canv <- canvassing$respondent_id[duplicated(canvassing$respondent_id)]
if (length(dupes_canv) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes_canv)}}")
}

agreed <- !is.na(canvassing$continue_full) & canvassing$continue_full == "Yes - continuing"
battery_answered <- !is.na(canvassing$q1_child_now)
funnel_mismatch <- if_else(
  canvassing$response_status == "complete",
  xor(agreed, battery_answered),
  battery_answered & !agreed
)
if (any(funnel_mismatch)) {
  cli::cli_warn("Battery answered/skipped against the doorstep funnel for
                 {sum(funnel_mismatch)} respondent{?s}")
}
```

How contacts move through the doorstep funnel. Each row is one combination of
answered the door, reported a connection to care, agreed to continue, and
pathway.

```{r}
#| label: canvassing-funnel-count
#| echo: false
count(canvassing, response_status, responded, care_connection, continue_full, pathway,
      name = "contacts") |>
  tbl("Canvassing contacts by funnel stage and pathway")
```

```{r}
#| label: canvassing-write
#| include: false
write_rds(canvassing, here("data/processed/canvassing_clean.rds"))
write_csv(canvassing, here("data/processed/canvassing_clean.csv"))
```

## Data dictionary

The column names and answer lists above already describe every column in the
clean files. This writes them out as a single table — one row per column (or
per level of a single-answer question), with the question and answer text it
came from — so anyone reading the clean files can look up what a column means
without opening this program.

```{r}
#| label: dictionary
question_text <- c(col_map, col_map_canv[setdiff(names(col_map_canv), names(col_map))])
all_multiselect <- c(multiselect_options, list(doorstep = doorstep_options))

dict_options <- all_multiselect |>
  imap(\(opts, q) tibble(
    column   = str_c(q, "_", c(names(opts), "other", "other_text")),
    question = unname(question_text[q]),
    # aliased options (form revisions) list every accepted wording
    option   = c(unname(map_chr(opts, str_flatten, collapse = " [or] ")),
                 "Other (any response not in the structured options)",
                 "Free text captured from other responses")
  )) |>
  list_rbind()

dict_factors <- list(survey, canvassing) |>
  map(\(data) {
    data |>
      select(where(is.factor)) |>
      as.list() |>
      imap(\(v, nm) tibble(
        column   = nm,
        question = unname(question_text[nm]),
        option   = levels(v)
      )) |>
      list_rbind()
  }) |>
  list_rbind() |>
  distinct()

dictionary <- bind_rows(dict_options, dict_factors) |>
  mutate(question = if_else(
    column == "pathway",
    "Derived from q1 (priority: current > past > future > observer > none)",
    question
  ))

write_csv(dictionary, here("data/processed/data-dictionary.csv"))
```

Search or page through the dictionary here; the same table is written to
`data/processed/data-dictionary.csv`.

```{r}
#| label: dictionary-show
#| echo: false
dictionary |>
  gt() |>
  tab_options(table.font.size = px(13)) |>
  opt_interactive(use_search = TRUE, use_page_size_select = TRUE, page_size_default = 15)
```
