---
title: "Cleaning the survey and canvassing data"
author: "Manuel Alcalá Kovalski, Urban Institute"
date: today
execute:
  warning: true
  message: false
---

The coalition delivers raw exports as Excel files to a Box folder. This
program reads a local copy of that folder, cleans it, and writes four files:

- `survey_clean.csv`: one row per survey respondent
- `canvassing_clean.csv`: one row per canvassing contact (original form)
- `canvassing_revised_clean.csv`: one row per contact on the revised form (August 2026 on)
- `data-dictionary.csv`: every column in those files, with the question and
  answer text it came from and, for single-answer questions, the order of the
  answers

## The survey in brief

The survey starts with a question about a person's connection to care. The
answer decides which of five sets of questions they see next, named for who
was asked:

- `current` (a): giving or receiving care now
- `past` (c): cared for someone in the past
- `future` (b): expects to give or need care in the next five years
- `observer` (d): knows someone who receives or needs care
- `none`: none of these, so no further questions until ideal care

They are listed in routing order: a person who ticks answers from more than
one group is meant to see the first set that applies.

Everyone then answers the same question about ideal care and the demographic
questions.

```{mermaid}
%%| label: fig-routing
%%| fig-cap: "How respondents move through the survey, in routing order: the first pathway that applies wins."
flowchart TD
  q1["q1 · Connection to care<br/>(choose all that apply)"]
  q1 -->|"caring for someone now, paid provider,<br/>or needs / receives care"| A
  q1 -->|"cared for someone in the past"| C
  q1 -->|"expects to give or need care"| B
  q1 -->|"knows someone who receives care"| D
  q1 -->|"none of these"| E

  A["current (a)<br/>q2a challenges · q3a supports · q4a who helps"]
  C["past (c)<br/>q2c · q3c · q4c"]
  B["future (b)<br/>q2b · q3b"]
  D["observer (d)<br/>q2d · q3d · q4d"]
  E["none"]

  A --> q6
  C --> q6
  B --> q6
  D --> q6
  E --> q6
  q6["q6 · Ideal care (everyone)"] --> demo["Demographics<br/>ZIP · household · income · age · gender<br/>race · disability · political views"]
```

::: {.callout-warning title="One routing exception, not yet confirmed"}
The data mostly bear the routing rule out: a person eligible for both `past` and
`future` was shown only the `past` questions. The exception is paid care
providers who also cared for someone in the past, who were shown both the
`current` and the `past` questions. We have not yet confirmed with Caring
Across Generations whether that is intended. The cleaning assigns those
respondents to `current`.
:::

Column names in the clean files follow this structure. For the survey
questions, the number is the topic and the letter is the pathway: the topics
are 1, connection to care; 2, challenges; 3, supports that would help; 4, who
helps; and 6, ideal care. So `q2a` is the challenges question asked of
current caregivers, and `q3d` is the supports question asked of people who
know someone receiving care. Demographic columns get plain names such as
`age` and `hh_income`. The [Variables page](/variables.qmd) lists every
column with the question and answer text behind it.

## The steps at a glance

1. **Find the files.** Use the team's data log to identify what each
   delivered file is; clean the survey and canvassing exports, and set aside
   test deliveries, open-text exports, and the pre-launch poll. Every file
   must be accounted for.
2. **Rename the columns** from full question text to short names (`q1`,
   `q2a`, `age`, ...), checking that nothing is missing or unexpected.
3. **Check for personal information.** Address columns must be empty.
4. **Split the choose-all-that-apply questions** into one TRUE/FALSE column
   per answer, keeping "Other" text word for word.
5. **Put single-answer questions in order** (income, age, and so on).
6. **Work out which set of questions each person was shown** (their
   "pathway") from their first answer, and **check** that they answered
   those questions and no others.
7. **Tidy the rest**: IDs, timestamps, ZIP codes. Confirm no respondent
   appears twice.
8. **Do the same for canvassing**, reusing the survey's answer lists: once
   for the original form and once for the revised form introduced in August,
   which asks different questions.
9. **Write the clean files** and a data dictionary describing every column.

::: {.callout-warning title="Warnings in this document are expected"}
Each step compares the delivery with what we have seen before. A renamed
column, an answer option not on our list, or a file the data log does not
describe stops the program with a message saying what changed. When that
happens, the fix is to update the list or the rule, not to skip the check.

Some checks warn instead of stopping. Those warnings appear below marked with
`!`. The known ones (paid care providers shown two sets of questions, ZIP+4
entries, one canvassing contact who answered questions without agreeing to
continue) are expected. A new one is the signal to investigate.
:::

::: {.callout-note title="The live form does not match the questionnaire PDF"}
The wording in the live form differs from the official question documents in
many places, and the export format itself has changed between deliveries
(how answers are separated, punctuation, and the wording of nine questions).
Answer and question text in this program comes from what we actually observe
in the exports, not from the PDF, and we keep a running record of how the
form really behaves as deliveries arrive.
:::

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(readxl)
library(here)
library(reactable)
library(urbnthemes)

set_urbn_defaults(style = "print")
source(here("scripts/00_utils.R"))
```

```{=html}
<style>
.reactable { margin: 0.75rem 0 1.75rem; }
.table-title { margin: 1.25rem 0 0.4rem; }
</style>
```

```{r}
#| label: table-helper
#| include: false
# one look for every table in this document: sortable, paged past `page_size`
# rows, and scrolling sideways rather than wrapping when there are many columns
tbl <- function(data, title = NULL, page_size = 5, searchable = FALSE, min_width = 90,
                wrap = TRUE, ...) {
  table <- reactable(
    data,
    pagination = nrow(data) > page_size, defaultPageSize = page_size,
    showPageSizeOptions = FALSE, searchable = searchable,
    striped = TRUE, resizable = TRUE, wrap = wrap,
    defaultColDef = colDef(minWidth = min_width),
    ...,
    theme = reactableTheme(
      cellPadding = "8px 12px",
      headerStyle = list(background = "#f3f4f6", fontWeight = 600),
      borderColor = "#e5e7eb"
    ),
    style = list(fontSize = "14px")
  )
  if (is.null(title)) return(table)
  htmltools::tagList(
    htmltools::tags$p(htmltools::tags$strong(title), class = "table-title"),
    table
  )
}
```

## Find the files to clean

The research team's data log on Box records what every delivered file is:
which tool (survey or canvassing), which link or form (`f1`, `f2`, `c1`, ...),
whether it holds closed-ended or open-text answers, complete or partial
responses, whether it was a pre-launch test, the dates it covers, and the
range of respondent IDs in it. This program uses that log to decide what each
file is. The file name is used only as a cross-check, and a disagreement
between the two is flagged.

This program cleans the closed-ended survey and canvassing exports. It leaves
out the pre-launch test deliveries, and it does not read the open-text
exports, the pre-launch poll, or the interview transcripts, which go to the
text-analysis pipeline. Every file in the folder is either cleaned here or
deliberately left out for one of those reasons, so a new kind of delivery
cannot slip through unnoticed.

```{r}
#| label: read
raw_dir <- here("data/raw/Raw data backups")

deliveries <- read_csv(here("data/raw/deliveries.csv"), show_col_types = FALSE)

mirror_files <- list.files(raw_dir) |>
  str_subset("~\\$", negate = TRUE) |> # Excel lock files, present while a workbook is open
  str_subset("DATA LOG", negate = TRUE)

# every mirrored file must be described in the delivery table
undescribed <- setdiff(mirror_files, deliveries$file)
if (length(undescribed) > 0) {
  cli::cli_abort(c(
    "{length(undescribed)} file{?s} in the mirror not described in deliveries.csv — re-run 00_ingest-raw.R:",
    set_names(undescribed, rep("x", length(undescribed)))
  ))
}

real_closed <- deliveries |>
  filter(folder == "Raw data backups", !test, content == "closed")

survey_files <- real_closed |> filter(tool == "survey") |> pull(file)
# the canvassing form was revised in Aug 2026 (forms numbered c1, c2, ...);
# the earlier form's files say f1. The two ask different questions and are
# cleaned separately.
canvassing_files         <- real_closed |> filter(tool == "canvass", str_starts(form, "f")) |> pull(file)
canvassing_revised_files <- real_closed |> filter(tool == "canvass", str_starts(form, "c")) |> pull(file)

set_aside <- deliveries |>
  filter(!file %in% c(survey_files, canvassing_files, canvassing_revised_files)) |>
  mutate(why = case_when(
    test                ~ "pre-launch test delivery",
    tool == "interview" ~ "interview transcript — text-analysis pipeline",
    tool == "poll"      ~ "pre-launch One Question Poll — text-analysis pipeline",
    content == "open"   ~ "open-text export — text-analysis pipeline",
    TRUE                ~ NA_character_
  ))
if (any(is.na(set_aside$why))) {
  cli::cli_abort(c(
    "Deliver{?y/ies} neither cleaned here nor covered by a set-aside rule:",
    set_names(set_aside$file[is.na(set_aside$why)], rep("x", sum(is.na(set_aside$why))))
  ))
}

# what the log says about each file, carried onto every row it contributes
delivery_info <- deliveries |>
  transmute(delivery = file, form_version = form, response_status = status)

survey_files <- file.path(raw_dir, survey_files)
canvassing_files <- file.path(raw_dir, canvassing_files)
canvassing_revised_files <- file.path(raw_dir, canvassing_revised_files)
```

The files cleaned in this run, as the data log describes them, are listed
under the fold.

::: {.callout-note collapse="true" title="Files cleaned in this run"}
```{r}
#| label: files-cleaned
#| echo: false
deliveries |>
  filter(file %in% basename(c(survey_files, canvassing_files, canvassing_revised_files))) |>
  transmute(
    file, tool, form, status, delivered,
    covers = if_else(is.na(collection_start), "not recorded",
                     str_c(collection_start, " to ", collection_end))
  ) |>
  arrange(desc(delivered), file) |>
  tbl(min_width = 110)
```
:::

## Survey

### Give the columns short names

In the export, each column is named with the full question text. We map those
to the short names described in "The survey in brief" above (`q2a`, `q3d`,
`age`, ...).

The mapping is checked in both directions: a column we have no name for, or
a column we expect but cannot find, stops the program. Since August the
exports carry new wording for nine of the questions; both wordings are listed
for those questions and lead to the same column. The three below show the
kind of change involved.

```{r}
#| label: rename
# where a question has two wordings, the first is the original form (June
# 2026) and the second the revised wording that appears in exports from
# August 2026; both name the same question
col_map <- list(
  submitted_at  = "Time",
  q1            = c("We would like to know more about your connection to care, if any. Choose all that apply to you.",
                    "Care and caregiving look different for everyone. Choose all that apply to you."),
  q2a           = c("What challenges, if any, have you experienced with providing, finding, or receiving care? (Choose all that apply)",
                    "What's been hard? (Choose all that apply)"),
  q3a           = c("What care support would make a difference for you or people around you? (Choose all that apply)",
                    "What help would make a difference for you or the people you care for? (Choose all that apply)"),
  q4a           = "What or who helps you most with care? (Choose your top 3)",
  q2b           = c("When you think about the care you, a family member, or friend might need 5 years from now, what challenges, if any, concern you most? (Choose all that apply)",
                    "What do you think will be hard in the future? (Choose all that apply)"),
  q3b           = c("What types of care support and services do you think will matter most for you or people you know? (Choose all that apply)",
                    "What help would make a difference? (Choose all that apply)"),
  q2c           = c("When you provided or received care what challenges, if any, did you experience? (Choose all that apply)",
                    "What was hard? (Choose all that apply)"),
  q3c           = c("What care support would have made a difference for you or people around you? (Choose all that apply)",
                    "What help would have made difference for you or people around you? (Choose all that apply)"),
  q4c           = "What or who helped you most with care? (Choose your top 3)",
  q2d           = c("What care challenges, if any, do the people you know experience? (Choose all that apply.)",
                    "What’s been hard for the people you know? (Choose all that apply.)"),
  q3d           = c("What care support would make a difference for the people you know? (Choose all that apply)",
                    "What help would make a difference for the people you know? (Choose all that apply)"),
  q4d           = "What or who helps most with their care? (Choose your top 3)",
  q6            = "What would ideal care support and services look like for you? (Choose up to three)",
  address       = "Zipcode (Address)",
  address2      = "Zipcode (Address2)",
  city          = "Zipcode (City)",
  state         = "Zipcode (State)",
  zip           = "Zipcode (Zip)",
  hh_size       = "How many people live in your household? (Include yourself and anyone you share income and living expenses with, including kids)",
  hh_income     = "What is your household's approximate annual income? (Add up income from everyone living with you, related or not.)",
  age           = "What is your age?",
  gender        = "What is your gender? (Choose all that apply)",
  race          = "What is your race or ethnicity? (Choose all that apply)",
  disability    = "I identify as a person with a disability or chronic health condition.",
  political     = "How would you describe your political views?",
  respondent_id = "Unique ID"
)

form_varying <- c("address", "address2", "city", "state", "q4c")

# exports from Aug 2026 on label the timestamp column "Date submitted"
# instead of "Time"; same Excel serial values underneath
timestamp_aliases <- c("Time" = "Date submitted")

# the same export writes en-/em-dashes and curly apostrophes ("$25,000–$49,999",
# "Affordable — without ...", "I don’t") where every other export writes plain
# hyphens and straight quotes
normalize_punctuation <- \(x) str_replace_all(x, c("—" = "-", "–" = "-", "’" = "'", "‘" = "'"))

# Read one export: every cell as text (so IDs and ZIPs keep their digits),
# then normalize punctuation, blank out whitespace-only cells, map the headers
# to short names, and turn the timestamp into a date-time.
read_export <- function(path, col_map, optional = character()) {
  read_excel(path, col_types = "text") |>
    rename(any_of(timestamp_aliases)) |>
    mutate(across(everything(), \(cell) na_if(str_trim(normalize_punctuation(cell)), ""))) |>
    rename_validated(col_map, optional = optional) |>
    mutate(
      delivery = basename(path),
      # timezone assumed UTC; the platform's export timezone is unconfirmed
      submitted_at = excel_datetime(submitted_at)
    )
}

survey <- survey_files |>
  map(\(path) read_export(path, col_map, optional = form_varying)) |>
  list_rbind() |>
  left_join(delivery_info, by = join_by(delivery))
```

```{r}
#| label: wordings
#| echo: false
col_map |>
  keep(\(wordings) length(wordings) == 2) |>
  imap(\(wordings, column) tibble(column, `wording until July` = wordings[1], `wording from August` = wordings[2])) |>
  list_rbind() |>
  slice(1:3) |>
  tbl(min_width = 220)
```

The table below counts the responses in each file. `form_version` is the
link or form the response came through (the May launch link, the July Virtual
Day of Action link, postcards, and so on; all use the same questionnaire), and
`response_status` separates people who finished the survey from those who
stopped partway through. Both stay in the clean file so the analysis can
decide how to treat them.

```{r}
#| label: read-count
#| echo: false
count(survey, delivery, form_version, response_status, name = "responses") |>
  tbl("Responses in each file", columns = list(delivery = colDef(minWidth = 380)))
```

This is what the raw export looks like at this point: one row per person,
with every choose-all-that-apply answer packed into a single cell. Scroll
sideways for the rest of the columns; the two formats the export has used
(answers separated by vertical bars, and in the July–August complete file by
commas) are both visible in the challenges column.

```{r}
#| label: raw-sample
#| echo: false
rows_to_show <- 25 # raise to nrow(survey) to page through every row

sample_rows <- c(
  which(str_detect(survey$q2a, fixed("|")))[1],
  which(str_detect(survey$delivery, "20260710-20260812_f1_complete") & !is.na(survey$q2a))[1]
)

survey |>
  slice(c(sample_rows, setdiff(seq_len(rows_to_show), sample_rows))) |>
  tbl(min_width = 200, wrap = FALSE)
```

### Check for personal information

::: {.callout-important title="No personal information"}
The data shared with Urban must contain no personal information. ZIP code is
collected on purpose; street address is not. Some exports carry address
columns, which should always be empty. If one ever holds a value, the program
stops here so the file can be raised with the research team before anything
else happens.
:::

```{r}
#| label: pii-guard
# Stop if any of the named address columns holds a value; otherwise drop them.
assert_no_address <- function(data, address_columns) {
  address_columns <- intersect(address_columns, names(data))
  has_values <- data |>
    select(all_of(address_columns)) |>
    map_lgl(\(column) any(!is.na(column)))
  if (any(has_values)) {
    cli::cli_abort(c(
      "Address column{?s} {.val {names(has_values)[has_values]}} contain{?s/} values — possible personal information.",
      "x" = "Do not proceed. Flag the delivery to the research team."
    ))
  }
  select(data, -all_of(address_columns))
}

survey <- assert_no_address(survey, c("address", "address2"))
```

### Split the choose-all-that-apply questions into one column per answer

Most survey questions let people choose more than one answer. The export
stores all of a person's choices in a single cell, separated by vertical
bars, for example `"Care is unaffordable|Other: my own words"`. This step
splits each question into one TRUE/FALSE column per answer option. For the
challenges question `q2a` (care is unaffordable, hard to find care, not enough
time or flexibility, and so on), that means columns such as
`q2a_unaffordable` and `q2a_time_flex`.

One delivery (the July–August complete responses) separates the answers with
commas instead. That is trickier, because the answers themselves contain
commas, so for those cells the known answers are matched one at a time from
the start of the cell.

::: {.callout-tip title="How to read the TRUE/FALSE columns"}
Each answer column holds one of three values:

- `TRUE`: the person chose this answer
- `FALSE`: the person saw the question and did not choose this answer
- `NA`: the person never saw the question (the survey routed them past it) or
  left it blank

The share of people choosing an answer is the share of `TRUE` among the
non-missing values.
:::

The answer options for every question are listed in the code below. The
short name on the left becomes the column suffix.

```{r}
#| label: dictionaries
multiselect_options <- list(
  q1 = c(
    child_now      = "I care for a child or children now",
    aging_now      = "I care for an aging person now",
    disability_now = "I care for someone with a disability or chronic health condition now",
    paid_provider  = "I am a paid care provider. I offer support and services for kids, aging adults, and/or people with disabilities",
    need_care_now  = "I need care but am not receiving it",
    receives_care  = "I receive or have received care for a disability or chronic condition",
    cared_past     = "I have cared for an aging adult, child, or someone with a disability or chronic health condition in the past",
    expect_future  = "I am not providing or receiving care right now, but I expect to in the next 5 years",
    observer       = "None of these apply to me right now, but I know someone who receives or needs care",
    none           = "None of these apply to me"
  ),
  q2a = list(
    unaffordable = "Care is unaffordable",
    finding      = c(
      "It is hard to find care that's available, nearby, or right for my family or me",
      # Aug 19 2026 revision — seeded from the questionnaire PDF, not yet observed in an export
      "It's hard to find care that's available, nearby, or right for me or my family"
    ),
    time_flex    = "I don't have enough time or flexibility",
    help         = "I don't have enough help or support",
    figure_out   = "It is hard to figure out what help exists or how to get it",
    navigate     = "It is hard to navigate existing care programs",
    no_say       = "I don't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3a = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    everyday_support   = "Everyday support like meals and transportation",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4a = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, child care, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid for using federal, state, or county dollars",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I do not receive help"
  ),
  q2b = c(
    unaffordable = "Care that's unaffordable",
    finding      = "Difficulty with finding care that's available, nearby, or right for me or my family",
    time_flex    = "Not having time or flexibility",
    help         = "Not enough help or support with care",
    figure_out   = "Difficulty knowing what help exists or how to get it",
    navigate     = "Difficulty navigating existing care programs",
    no_say       = "Not having a say in how care works for my family or me",
    job          = "Care could affect my job or career"
  ),
  q3b = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q2c = c(
    unaffordable = "Care was unaffordable",
    finding      = "It was hard to find care that was available, nearby, or right for my family and me",
    time_flex    = "I didn't have enough time or flexibility",
    help         = "I didn't have enough help or support",
    figure_out   = "It was hard to figure out what help existed or how to get it",
    navigate     = "It was hard to navigate existing care programs",
    no_say       = "I didn't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3c = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4c = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  q2d = c(
    unaffordable = "Unaffordable care",
    finding      = "It was hard to find care that's available, nearby, or right for them",
    time_flex    = "They don't have enough time or flexibility",
    help         = "They are doing this without enough help or support",
    figure_out   = "It is hard for them to figure out what help exists or how to get it",
    navigate     = "It is hard for them to navigate existing care programs",
    no_say       = "They don't have a say in how care works for them",
    job          = "Care has affected their jobs or careers"
  ),
  q3d = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4d = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    # in the instrument (per the Aug 19 2026 PDF, wording as in q4c) but unobserved so far
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  # Aug 19 2026 variants are seeded from the questionnaire PDF with the
  # formatting pattern of observed exports (plain hyphens, not the PDF's
  # em-dashes) — confirm against the first export that carries them
  q6 = list(
    there_when_needed = "There when you need it - at every stage of life, no matter your job or family situation",
    easy_to_use       = c(
      "Easy to find and use - without complicated steps or barriers",
      "Easy to find and use - without confusing steps or barriers"          # Aug 19 2026
    ),
    affordable        = "Affordable - without causing personal financial strain",
    in_community      = "Available in your community - not far away or out of reach",
    flexible          = "Flexible - fitting real schedules, needs, and changing circumstances",
    reliable          = "Reliable - something you can count on over time",
    clear             = "Clear and transparent - you know what's available and how to get it",
    good_jobs         = "Good jobs for care providers - with fair pay and stability",
    healthy_dev       = c(
      "Support healthy development, well-being, and independence",
      "Supports healthy development, well-being, and independence"          # Aug 19 2026
    ),
    lived_experience  = c(
      "Shaped by people with experience giving, receiving, or needing care",
      "Shaped by people with lived care experience"                         # Aug 19 2026
    ),
    secure            = "Keep people who rely on care feeling secure, healthy, and thriving"
  ),
  gender = c(
    woman             = "Woman",
    man               = "Man",
    nonbinary         = "Non-binary",
    transgender       = "Transgender",
    prefer_not_to_say = "Prefer not to say"
  ),
  race = c(
    aian              = "American Indian or Alaska Native",
    asian             = "Asian or Asian American",
    black             = "Black or African American",
    hispanic          = "Hispanic, Latino, or Latina",
    mena              = "Middle Eastern or North African",
    nhpi              = "Native Hawaiian or Pacific Islander",
    white             = "White",
    multiracial       = "Multiracial",
    prefer_not_to_say = "Prefer not to say"
  )
)

# the challenges cells as delivered, kept for the worked example below
raw_q2a <- survey$q2a

survey <- reduce(
  names(multiselect_options),
  \(data, col) encode_multiselect(data, col, multiselect_options[[col]]),
  .init = survey
)
```

Every choose-all-that-apply question ends with an "Other" checkbox and a
text box next to it, where people can type an answer of their own. In the
export, whatever they typed lands in the same packed cell as their ticked
answers. It can appear in three ways, depending on what the person did:

- they ticked "Other" and typed something: the cell contains `Other:`
  followed by their words
- they ticked "Other" but typed nothing: the cell contains just the word
  `Other`
- occasionally the export drops the label: the cell contains their words on
  their own, sometimes followed by a stray `true` (the checkbox's own value)

The cleaning treats all three the same way. Anything in a cell that is not
one of the standard answers is taken as an "Other" answer: the question's
`_other` column becomes `TRUE`, and the words, if there are any, are kept
unchanged in its `_other_text` column. Those words go to the text analysis.

The table below shows four real respondents, one for each case: standard
answers only, "Other" with typed words, "Other" ticked with nothing typed,
and typed words with no label. The first column is the challenges cell
exactly as delivered; the other columns are what the cleaning makes of it.

```{r}
#| label: other-example
#| echo: false
example_rows <- c(
  standard   = which(str_detect(raw_q2a, fixed("|")) & !survey$q2a_other & nchar(raw_q2a) < 90)[1],
  labelled   = which(str_detect(raw_q2a, "\\|Other: ") & nchar(raw_q2a) < 90)[1],
  bare       = which(raw_q2a == "Other")[1],
  unlabelled = which(survey$q2a_other & !str_detect(raw_q2a, "Other") & nchar(raw_q2a) < 60)[1]
) |>
  na.omit()

tibble(`export cell` = raw_q2a[example_rows]) |>
  bind_cols(survey[example_rows, c("q2a_unaffordable", "q2a_time_flex", "q2a_other", "q2a_other_text")]) |>
  tbl(min_width = 120)
```

Two safeguards sit inside this step. If the same "other" text appears from
three or more people, the program warns: repeated text usually means a new
answer option was added to the form and needs to be added to our list. And
when the form's wording for an answer changes, both the old and the new
wording are listed for that answer, so responses from before and after the
change land in the same column.


::: {.callout-note title="q6 asks for up to three answers, but the form does not enforce it"}
The ideal-care question `q6` asks people to choose up to three answers. The
live form does not enforce that limit, and many people choose more. We keep
every selection as given; the chartbook reports `q6` both ways. The count
below shows how many answers people actually chose.
:::

```{r}
#| label: q6-limit
#| echo: false
survey |>
  filter(!is.na(q6_affordable)) |>
  mutate(
    n_chosen = rowSums(across(starts_with("q6_") & where(is.logical) & !q6_other)),
    answers  = if_else(n_chosen > 3, "more than 3", as.character(n_chosen)) |>
      factor(levels = c("1", "2", "3", "more than 3"))
  ) |>
  count(answers, name = "respondents") |>
  mutate(share = scales::percent(respondents / sum(respondents), accuracy = 1)) |>
  tbl("Number of q6 answers chosen")
```

### Put single-answer questions in order

The demographic questions with one answer each become ordered categories
(household size from smallest to largest, income from lowest to highest, and
so on) so that tables and charts sort sensibly. The set of answers comes
from the questionnaire; an answer outside it stops the program.

```{r}
#| label: factors
survey <- survey |>
  mutate(
    # the export writes household size as "2.0", "3.0", ...
    hh_size = to_factor(str_remove(hh_size, "\\.0$"), c(
      "1, just myself", "2", "3", "4", "5", "6", "7 or more",
      "Prefer not to say"
    )),
    hh_income = to_factor(hh_income, c(
      "Under $25,000", "$25,000-$49,999", "$50,000-$74,999",
      "$75,000-$99,999", "$100,000-$149,999", "$150,000-$174,999",
      "$175,000-$199,999", "$200,000-$249,999", "$250,000-$299,999",
      "$300,000 or more", "Prefer not to say"
    )),
    age = to_factor(age, c(
      "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older",
      "Prefer not to say"
    )),
    disability = to_factor(disability, c(
      "Yes", "No", "Not sure", "Prefer not to say"
    )),
    political = to_factor(political, c(
      "Very liberal", "Somewhat liberal", "Moderate/Middle of the road",
      "Somewhat conservative", "Very conservative", "Don't know",
      "Prefer not to say"
    ))
  )
```

### Work out each person's pathway

The first question decides which set of questions a person sees (the five
pathways are described in "The survey in brief" above). The cleaning assigns
each person the first pathway that applies, in the questionnaire's order.

```{r}
#| label: pathway
survey <- derive_pathway(survey)
```

Respondents on each pathway. "No first answer" means the person stopped
before answering the first question.

```{r}
#| label: pathway-count
#| echo: false
survey |>
  mutate(pathway = fct_na_value_to_level(pathway, "no first answer")) |>
  count(pathway) |>
  ggplot(aes(n, fct_rev(pathway))) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 3.5) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Survey respondents by pathway", x = "Respondents", y = NULL)
```

### Check that people answered the questions for their pathway

Each pathway's questions should be answered by the people on that pathway and
nobody else. Partial responses may stop early, so for them only answers from
outside their pathway count against the check.

::: {.callout-note title="The known exception shows up here"}
Paid care providers who also cared for someone in the past were shown two
sets of questions (see "The survey in brief"). They are assigned to
`current`, so their answers to the `past` questions appear in the table below
as answers from outside the pathway. That is expected until Caring Across
Generations confirms the routing.
:::

```{r}
#| label: skip-checks
skip_rules <- tribble(
  ~question, ~indicator,             ~expected_pathway,
  "q2a",     "q2a_unaffordable",     "current",
  "q3a",     "q3a_backup",           "current",
  "q4a",     "q4a_family_friends",   "current",
  "q2b",     "q2b_job",              "future",
  "q3b",     "q3b_financial",        "future",
  "q2c",     "q2c_job",              "past",
  "q3c",     "q3c_backup",           "past",
  "q4c",     "q4c_family_friends",   "past",
  "q2d",     "q2d_unaffordable",     "observer",
  "q3d",     "q3d_backup",           "observer",
  "q4d",     "q4d_family_friends",   "observer"
)

# q6 is shown to every pathway, so it has no row in skip_rules — instead a
# complete response that skipped it entirely signals routing/export drift
q6_skipped <- survey$response_status == "complete" & is.na(survey$q6_affordable)
if (any(q6_skipped)) {
  cli::cli_warn("q6: {sum(q6_skipped)} complete response{?s} with q6 entirely missing")
}

# one row per pathway question: who answered it without being on the pathway,
# and which completes on the pathway skipped it
skip_checks <- skip_rules |>
  pmap(\(question, indicator, expected_pathway) {
    answered   <- !is.na(survey[[indicator]])
    on_pathway <- survey$pathway %in% expected_pathway
    complete   <- survey$response_status == "complete"
    tibble(
      question, pathway = expected_pathway,
      `answered, not on pathway`  = sum(answered & !on_pathway),
      `on pathway, skipped (completes)` = sum(complete & on_pathway & !answered)
    )
  }) |>
  list_rbind()

off_pathway <- sum(skip_checks$`answered, not on pathway`)
skipped     <- sum(skip_checks$`on pathway, skipped (completes)`)
if (off_pathway + skipped > 0) {
  cli::cli_warn("Routing: {off_pathway} answer{?s} from outside the pathway and {skipped} skipped on the pathway (table below)")
}
```

```{r}
#| label: skip-checks-table
#| echo: false
tbl(skip_checks, "Answers against the survey's routing", page_size = nrow(skip_checks))
```

### Tidy the remaining columns

Two small fixes:

- **Respondent IDs** sometimes export in scientific notation, so they are
  rewritten as plain digits, which also lets us confirm that nobody appears
  twice.
- **ZIP codes** pick up a stray `.0` from the export and lose their leading
  zeros, so both are repaired. Any ZIP that still is not five digits is
  flagged.

```{r}
#| label: other-cols
survey <- survey |>
  mutate(
    respondent_id = normalize_id(respondent_id),
    zip = str_pad(str_remove(zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

# after normalization so ID collisions across export formats would surface
dupes <- survey$respondent_id[duplicated(survey$respondent_id)]
if (length(dupes) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes)}}")
}

bad_zip <- survey |> filter(!is.na(zip), !str_detect(zip, "^\\d{5}$"))
if (nrow(bad_zip) > 0) {
  cli::cli_warn("{nrow(bad_zip)} row{?s} with a ZIP that is not five digits (ZIP+4 entries, kept as typed)")
}
```

```{r}
#| label: zip-missing
#| include: false
zip_missing <- survey |>
  summarise(share_missing = mean(is.na(zip)), .by = response_status) |>
  deframe()
if (zip_missing[["complete"]] > 0) {
  cli::cli_warn("{scales::percent(zip_missing[['complete']], accuracy = 0.1)} of complete responses have no ZIP code")
}
```

Every complete response has a ZIP code. Among partial responses,
`r scales::percent(zip_missing[["partial"]], accuracy = 1)` do not, which is
expected: ZIP is the first demographic question, and many partials stop
before reaching it.

### Check each file against the data log

The research team's data log records the dates each file covers and the first and last
respondent ID in it. Each file's rows are compared with that: submission
dates should fall inside the logged window and IDs inside the logged range. A
mismatch means the wrong file was delivered or logged, which nothing else
would catch.

```{r}
#| label: log-check
survey_checks <- check_against_log(survey, deliveries, "Survey")
```

```{r}
#| label: log-check-table
#| echo: false
survey_checks |>
  transmute(
    delivery, rows,
    `rows cover` = str_c(first_day, " to ", last_day),
    `log says` = if_else(is.na(collection_start), "not recorded", str_c(collection_start, " to ", collection_end)),
    `IDs in range` = if_else(is.na(first_id) | is.na(last_id), "not recorded", if_else(ids_ok, "yes", "NO"))
  ) |>
  tbl("Survey files against the data log")
```

```{r}
#| label: write
#| include: false
write_csv(survey, here("data/processed/survey_clean.csv"))
```

The clean survey file has one row per respondent and
`r ncol(survey)` columns. Most are the TRUE/FALSE answer columns; the rest
identify the respondent and delivery or hold the single-answer demographics.

```{r}
#| label: survey-structure
#| echo: false
tibble(column = names(survey), type = map_chr(survey, \(x) class(x)[1])) |>
  mutate(
    group = case_when(
      str_detect(column, "^(q\\d[a-d]?|gender|race)_") ~ str_extract(column, "^[^_]+"),
      TRUE ~ "respondent / delivery"
    ),
    type = case_match(type, "logical" ~ "TRUE/FALSE", "factor" ~ "ordered category",
                      "character" ~ "text", "POSIXct" ~ "date-time", .default = type)
  ) |>
  summarise(columns = n(), types = str_flatten(unique(type), ", "),
            examples = str_flatten(head(column, 3), ", "), .by = group) |>
  tbl("Columns in the clean survey file")
```

## Canvassing

The canvassing form is used at people's doors. It starts with a few questions
for the canvasser (where they are, how the conversation is happening, whether
the person answered), a short screener, and a quick-pick list of care
challenges. People who agree to continue then get the same question set as
the survey, word for word, so the survey's column names and answer lists are
reused. There are no demographic questions and no respondent ZIP; the only
geography is the canvasser's ZIP.

As with the survey, the `f0` test delivery is left out, and complete and
partial responses are combined and labelled.

```{r}
#| label: canvassing-read
#| echo: false
basename(canvassing_files)
```

```{r}
#| label: canvassing-rename
battery <- col_map[c(
  "q1", "q2a", "q3a", "q4a", "q2b", "q3b", "q2c", "q3c", "q4c",
  "q2d", "q3d", "q4d", "q6"
)]

col_map_canv <- c(
  submitted_at        = "Time",
  canvasser_zip       = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Zip)",
  canvass_mode        = "For Canvasser: How are you conducting this conversation?",
  responded           = "For Canvasser: Did the person respond and agree to answer questions?  If no, share the information card or the careconversations.com website and submit.",
  care_connection     = "Have you provided, received care yourself, expected to in the future or do you know someone who is? This could be giving or receiving care for an aging family member, a friend, a child, or someone with a disability no matter their age. It also includes you.",
  doorstep             = "What challenges, if any, have you or others faced — or do you expect to face — when it comes to giving, finding, or getting care? (For canvasser: Check all that apply. If selecting 'No Challenges,' leave everything else unchecked.)",
  continue_full       = "We're talking with people across the country about their experiences with care — what's working and what families need.  Can you take 5-minutes to share more?",
  battery,
  respondent_id       = "Unique ID"
)

canvassing <- canvassing_files |>
  map(\(path) read_export(path, col_map_canv)) |>
  list_rbind() |>
  left_join(delivery_info, by = join_by(delivery))
```

```{r}
#| label: canvassing-count
#| echo: false
count(canvassing, delivery, form_version, response_status, name = "responses") |>
  tbl("Responses in each canvassing file", columns = list(delivery = colDef(minWidth = 380)))
```

The doorstep challenge list uses short labels of its own. Everything else
reuses the survey answer lists.

```{r}
#| label: canvassing-encode
doorstep_options <- c(
  cost                = "Cost",
  time                = "Time",
  location            = "Location",
  right_fit           = "Right fit",
  lack_support        = "Lack of support",
  no_control          = "No control",
  worker_availability = "Worker availability",
  barriers_access     = "Barriers to access",
  no_challenges       = "No challenges",
  # added to the form between f0 and f1 (surfaced by the repeat-drift alarm)
  availability        = "Availability",
  figure_out          = "Hard to figure out"
)

canv_multiselect <- c(
  list(doorstep = doorstep_options),
  multiselect_options[c(
    "q1", "q2a", "q3a", "q4a", "q2b", "q3b",
    "q2c", "q3c", "q4c", "q2d", "q3d", "q4d", "q6"
  )]
)

canvassing <- reduce(
  names(canv_multiselect),
  \(data, col) encode_multiselect(data, col, canv_multiselect[[col]]),
  .init = canvassing
) |>
  derive_pathway()
```

The same tidying as the survey, plus one check specific to canvassing: the
full question set should be answered by exactly the people who agreed to
continue. As before, partials may stop early but should not answer questions
without agreeing.

```{r}
#| label: canvassing-clean
canvassing <- canvassing |>
  mutate(
    respondent_id   = normalize_id(respondent_id),
    canvass_mode    = to_factor(canvass_mode, c("Door-to-door", "Phone", "Other In Person")),
    responded       = to_factor(responded, c("Yes - continue with questions", "No - not available")),
    care_connection = to_factor(care_connection, c("Yes", "No")),
    continue_full   = to_factor(continue_full, c("Yes - continuing", "No - not continuing")),
    # same float artifact as the survey ZIPs
    canvasser_zip   = str_pad(str_remove(canvasser_zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

dupes_canv <- canvassing$respondent_id[duplicated(canvassing$respondent_id)]
if (length(dupes_canv) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes_canv)}}")
}

canvassing_checks <- check_against_log(canvassing, deliveries, "Canvassing")

agreed <- !is.na(canvassing$continue_full) & canvassing$continue_full == "Yes - continuing"
battery_answered <- !is.na(canvassing$q1_child_now)
funnel_mismatch <- if_else(
  canvassing$response_status == "complete",
  xor(agreed, battery_answered),
  battery_answered & !agreed
)
if (any(funnel_mismatch)) {
  cli::cli_warn("Battery answered/skipped against the doorstep funnel for
                 {sum(funnel_mismatch)} respondent{?s}")
}
```

How contacts move through the doorstep funnel. Each row is one combination of
answered the door, reported a connection to care, agreed to continue, and
pathway.

```{r}
#| label: canvassing-funnel-count
#| echo: false
count(canvassing, response_status, responded, care_connection, continue_full, pathway,
      name = "contacts") |>
  tbl("Canvassing contacts by funnel stage and pathway")
```

```{r}
#| label: canvassing-write
#| include: false
write_csv(canvassing, here("data/processed/canvassing_clean.csv"))
```

## Canvassing, revised form

In August 2026 the coalition changed the canvassing form. The yes/no
"connection to care" question and the doorstep challenge checklist are gone;
in their place are two questions — "Do you need care or is there someone in
your life you help take care of?" and, as free text, "What's been hard about
that?" The revised form comes in three versions (`c1`, `c2`, `c3`) with
different sets of questions: `c1` keeps the canvasser fields and the full
survey questions, `c2` (used at the Daisy Chain event) has only the two new
questions, and `c3` has the canvasser fields and the two questions but not the
survey questions. Because the questions differ from the earlier form's, these
responses are cleaned into their own file and reported separately.

```{r}
#| label: canvassing-revised-read
#| echo: false
basename(canvassing_revised_files)
```

Every column any version can carry is mapped; the ones a given version lacks
are simply absent for its rows. The canvasser's street-address columns, which
`c3` carries, must be empty, as for the survey.

```{r}
#| label: canvassing-revised-rename
col_map_canv_revised <- c(
  submitted_at      = "Time",
  canvasser_address  = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Address)",
  canvasser_address2 = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Address2)",
  canvasser_city     = "For Canvasser: Enter the ZIP code of the place you are canvassing  (City)",
  canvasser_state    = "For Canvasser: Enter the ZIP code of the place you are canvassing  (State)",
  canvasser_zip     = col_map_canv[["canvasser_zip"]],
  canvass_mode      = col_map_canv[["canvass_mode"]],
  responded         = col_map_canv[["responded"]],
  screener          = "1. Do you need care or is there someone in your life you help take care of – a kid, a parent, a grandparent, or anyone else?",
  hard_text         = "2. What’s been hard about that?",
  continue_full     = col_map_canv[["continue_full"]],
  battery,
  respondent_id     = "Unique ID"
)
always_present <- c("submitted_at", "screener", "hard_text", "respondent_id")

canvassing_revised <- canvassing_revised_files |>
  map(\(path) read_export(path, col_map_canv_revised,
                          optional = setdiff(names(col_map_canv_revised), always_present))) |>
  list_rbind() |>
  left_join(delivery_info, by = join_by(delivery)) |>
  assert_no_address(c("canvasser_address", "canvasser_address2")) |>
  select(-any_of(c("canvasser_city", "canvasser_state")))
```

```{r}
#| label: canvassing-revised-count
#| echo: false
count(canvassing_revised, delivery, form_version, response_status, name = "responses") |>
  tbl("Responses in each revised-form file", columns = list(delivery = colDef(minWidth = 380)))
```

The survey questions, where a version carries them, are split into TRUE/FALSE
columns with the same answer lists as before. The free-text answers are kept
word for word for the text analysis and are not summarized here.

```{r}
#| label: canvassing-revised-clean
present <- intersect(names(canv_multiselect), names(canvassing_revised))
canvassing_revised <- reduce(
  present,
  \(data, col) encode_multiselect(data, col, canv_multiselect[[col]]),
  .init = canvassing_revised
)
if ("q1" %in% present) canvassing_revised <- derive_pathway(canvassing_revised)

canvassing_revised <- canvassing_revised |>
  mutate(
    respondent_id = normalize_id(respondent_id),
    screener      = to_factor(screener, c("Yes", "No")),
    across(any_of("canvass_mode"),    \(x) to_factor(x, c("Door-to-door", "Phone", "Other In Person"))),
    across(any_of("responded"),       \(x) to_factor(x, c("Yes - continue with questions", "No - not available"))),
    across(any_of("continue_full"),   \(x) to_factor(x, c("Yes - continuing", "No - not continuing"))),
    across(any_of("canvasser_zip"),   \(x) str_pad(str_remove(x, "\\.0$"), width = 5, pad = "0"))
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

dupes_rev <- canvassing_revised$respondent_id[duplicated(canvassing_revised$respondent_id)]
if (length(dupes_rev) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes_rev)}}")
}
revised_checks <- check_against_log(canvassing_revised, deliveries, "Canvassing (revised form)")
```

```{r}
#| label: canvassing-revised-summary
#| echo: false
canvassing_revised |>
  count(form_version, response_status, screener, wrote_text = !is.na(hard_text), name = "contacts") |>
  tbl("Revised-form contacts by version, status, screener answer, and free text")
```

```{r}
#| label: canvassing-revised-write
#| include: false
write_csv(canvassing_revised, here("data/processed/canvassing_revised_clean.csv"))
```

## Data dictionary

The column names and answer lists above already describe every column in the
clean files. This writes them out as a single table — one row per column (or
per level of a single-answer question), with the question and answer text it
came from — so anyone reading the clean files can look up what a column means
without opening this program.

```{r}
#| label: dictionary
# the dictionary shows the original wording of each question
question_text <- c(
  col_map,
  col_map_canv[setdiff(names(col_map_canv), names(col_map))],
  col_map_canv_revised[c("screener", "hard_text")]
) |>
  map_chr(\(x) x[1])
all_multiselect <- c(multiselect_options, list(doorstep = doorstep_options))

dict_options <- all_multiselect |>
  imap(\(opts, q) tibble(
    column   = str_c(q, "_", c(names(opts), "other", "other_text")),
    question = unname(question_text[q]),
    # aliased options (form revisions) list every accepted wording
    option   = c(unname(map_chr(opts, str_flatten, collapse = " [or] ")),
                 "Other (any response not in the structured options)",
                 "Free text captured from other responses")
  )) |>
  list_rbind()

dict_factors <- list(survey, canvassing, canvassing_revised) |>
  map(\(data) {
    data |>
      select(where(is.factor)) |>
      as.list() |>
      imap(\(v, nm) tibble(
        column   = nm,
        question = unname(question_text[nm]),
        option   = levels(v)
      )) |>
      list_rbind()
  }) |>
  list_rbind() |>
  distinct()

dictionary <- bind_rows(
  dict_options,
  dict_factors,
  tibble(column = "hard_text", question = unname(question_text["hard_text"]),
         option = "Free text, kept verbatim (revised canvassing form)")
) |>
  mutate(question = if_else(
    column == "pathway",
    "Derived from q1 (priority: current > past > future > observer > none)",
    question
  ))

write_csv(dictionary, here("data/processed/data-dictionary.csv"))
```

Search or page through the dictionary here; the same table is written to
`data/processed/data-dictionary.csv`.

```{r}
#| label: dictionary-show
#| echo: false
dictionary |>
  tbl(page_size = 15, searchable = TRUE, min_width = 140)
```
