---
title: "Clean test data"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
  message: false
---

Cleans the raw survey and canvassing exports in `data/raw/` into
analysis-ready files in `data/processed/`, plus a generated data dictionary.

The main transformation is one-hot encoding the multi-select questions: the raw
export stores multiple selections as a single pipe-delimited string, and we
break each question into one logical column per response option. `TRUE` means
the option was selected, `FALSE` means the respondent saw the question but did
not select it, and `NA` means the respondent never saw the question (skip
logic) or skipped it. Selection rates are just `mean(x, na.rm = TRUE)`.

Response option text comes from the observed data, not the survey PDF — the
live form's wording differs from the June 15 questionnaire document (e.g. the
PDF's "Caring for a child/children" appears in the data as "I care for a child
or children now"). Options not yet observed in the data are therefore not in
the dictionaries below; the encoding stops with an error listing any new value
so it can be added deliberately.

```{r}
#| label: setup
library(tidyverse)
library(readxl)
library(here)

source(here("scripts/00_utils.R"))
```

## Read raw survey data

```{r}
#| label: read
survey_files <- list.files(
  here("data/raw/survey"),
  pattern = "\\.xlsx$", full.names = TRUE
) |>
  str_subset("~\\$", negate = TRUE) # Excel lock files, present while a workbook is open

raw <- survey_files |>
  map(read_excel, col_types = "text") |>
  bind_rows() |>
  mutate(across(everything(), \(x) na_if(str_trim(x), ""))) # whitespace-only cells -> NA

dim(raw)
```

## Rename columns

Column names in the export are the full question text. We map them to short
IDs keyed to the questionnaire (q1, q2a, ..., q6). The mapping is validated in
both directions so a new export with added, dropped, or reworded columns fails
loudly instead of silently misaligning.

Note the test export does not include q4c or the open-text questions
(q5a/q5c/q5d, q7); they will need entries here when they appear.

```{r}
#| label: rename
col_map <- c(
  submitted_at  = "Time",
  q1            = "We would like to know more about your connection to care, if any. Choose all that apply to you.",
  q2a           = "What challenges, if any, have you experienced with providing, finding, or receiving care? (Choose all that apply)",
  q3a           = "What care support would make a difference for you or people around you? (Choose all that apply)",
  q4a           = "What or who helps you most with care? (Choose your top 3)",
  q2b           = "When you think about the care you, a family member, or friend might need 5 years from now, what challenges, if any, concern you most? (Choose all that apply)",
  q3b           = "What types of care support and services do you think will matter most for you or people you know? (Choose all that apply)",
  q2c           = "When you provided or received care what challenges, if any, did you experience? (Choose all that apply)",
  q3c           = "What care support would have made a difference for you or people around you? (Choose all that apply)",
  q2d           = "What care challenges, if any, do the people you know experience? (Choose all that apply.)",
  q3d           = "What care support would make a difference for the people you know? (Choose all that apply)",
  q4d           = "What or who helps most with their care? (Choose your top 3)",
  q6            = "What would ideal care support and services look like for you? (Choose up to three)",
  address       = "Zipcode (Address)",
  address2      = "Zipcode (Address2)",
  city          = "Zipcode (City)",
  state         = "Zipcode (State)",
  zip           = "Zipcode (Zip)",
  hh_size       = "How many people live in your household? (Include yourself and anyone you share income and living expenses with, including kids)",
  hh_income     = "What is your household's approximate annual income? (Add up income from everyone living with you, related or not.)",
  age           = "What is your age?",
  gender        = "What is your gender? (Choose all that apply)",
  race          = "What is your race or ethnicity? (Choose all that apply)",
  disability    = "I identify as a person with a disability or chronic health condition.",
  political     = "How would you describe your political views?",
  respondent_id = "Unique ID"
)

survey <- rename_validated(raw, col_map)
```

## PII guard

Data shared with Urban must contain no PII. ZIP is collected by design, but
the export includes street-address fields that should always be empty — stop
immediately if they ever are not.

```{r}
#| label: pii-guard
if (any(!is.na(survey$address) | !is.na(survey$address2))) {
  stop("Address fields are not empty — possible PII in the raw file. Do not proceed; flag to the team.")
}

survey <- select(survey, -address, -address2)
```

## One-hot encode multi-select questions

Selections arrive pipe-delimited in survey order, e.g.
`"Care is unaffordable|Other"`. Each question becomes one logical column per
option (`q2a_unaffordable`, `q2a_other`, ...), with every option the
respondent did not select set to `FALSE` and rows where the question was
unanswered set to `NA` across all of its option columns.

```{r}
#| label: encode-fn
encode_multiselect <- function(data, col, options) {
  # tokens are trimmed so "c | e" and "c|e" encode identically
  selections <- str_split(data[[col]], fixed("|")) |> map(str_trim)

  unknown <- setdiff(unique(unlist(selections)), c(options, NA))
  if (length(unknown) > 0) {
    cli::cli_abort(c(
      "New response option{?s} in {.field {col}} — add {?it/them} to its dictionary:",
      set_names(paste0('"', unknown, '"'), "x")
    ))
  }

  indicators <- options |>
    map(\(opt) map_lgl(selections, \(s) opt %in% s)) |>
    set_names(str_c(col, "_", names(options))) |>
    as_tibble() |>
    mutate(across(everything(), \(x) if_else(is.na(data[[col]]), NA, x)))

  data |>
    select(-all_of(col)) |>
    bind_cols(indicators)
}
```

A quick demonstration on a toy question with options `a`–`f`: a response of
`"c | e"` gets `TRUE` for c and e and `FALSE` for everything else, and an
unanswered row is `NA` throughout.

```{r}
#| label: encode-demo
tibble(q_demo = c("c | e", "a", NA)) |>
  encode_multiselect("q_demo", c(a = "a", b = "b", c = "c", d = "d", e = "e", f = "f"))
```

```{r}
#| label: dictionaries
multiselect_options <- list(
  q1 = c(
    child_now      = "I care for a child or children now",
    aging_now      = "I care for an aging person now",
    disability_now = "I care for someone with a disability or chronic health condition now",
    paid_provider  = "I am a paid care provider. I offer support and services for kids, aging adults, and/or people with disabilities",
    cared_past     = "I have cared for an aging adult, child, or someone with a disability or chronic health condition in the past",
    expect_future  = "I am not providing or receiving care right now, but I expect to in the next 5 years",
    none           = "None of these apply to me"
  ),
  q2a = c(
    unaffordable = "Care is unaffordable",
    time_flex    = "I don't have enough time or flexibility",
    figure_out   = "It is hard to figure out what help exists or how to get it",
    navigate     = "It is hard to navigate existing care programs",
    job          = "Care affected my job or career",
    other        = "Other"
  ),
  q3a = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    everyday_support   = "Everyday support like meals and transportation",
    other              = "Other"
  ),
  q4a = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, child care, or pre-k",
    community_religious  = "Community organizations or religious groups",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    no_help              = "I do not receive help",
    other                = "Other"
  ),
  q2b = c(
    finding = "Difficulty with finding care that's available, nearby, or right for me or my family",
    help    = "Not enough help or support with care",
    job     = "Care could affect my job or career"
  ),
  q3b = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses"
  ),
  q2c = c(
    unaffordable = "Care was unaffordable",
    finding      = "It was hard to find care that was available, nearby, or right for my family and me",
    time_flex    = "I didn't have enough time or flexibility",
    help         = "I didn't have enough help or support",
    figure_out   = "It was hard to figure out what help existed or how to get it",
    navigate     = "It was hard to navigate existing care programs",
    no_say       = "I didn't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3c = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care"
  ),
  q6 = c(
    there_when_needed = "There when you need it - at every stage of life, no matter your job or family situation",
    easy_to_use       = "Easy to find and use - without complicated steps or barriers",
    affordable        = "Affordable - without causing personal financial strain",
    in_community      = "Available in your community - not far away or out of reach",
    flexible          = "Flexible - fitting real schedules, needs, and changing circumstances",
    reliable          = "Reliable - something you can count on over time",
    clear             = "Clear and transparent - you know what's available and how to get it",
    good_jobs         = "Good jobs for care providers - with fair pay and stability",
    secure            = "Keep people who rely on care feeling secure, healthy, and thriving"
  ),
  gender = c(
    woman             = "Woman",
    man               = "Man",
    other             = "Other",
    prefer_not_to_say = "Prefer not to say"
  ),
  race = c(
    black             = "Black or African American",
    hispanic          = "Hispanic, Latino, or Latina",
    mena              = "Middle Eastern or North African",
    nhpi              = "Native Hawaiian or Pacific Islander",
    multiracial       = "Multiracial",
    other             = "Other",
    prefer_not_to_say = "Prefer not to say"
  )
)

survey <- reduce(
  names(multiselect_options),
  \(data, col) encode_multiselect(data, col, multiselect_options[[col]]),
  .init = survey
)
```

The pathway D questions (q2d, q3d, q4d) are entirely `NA` in the test data —
no observer-pathway respondents yet — so there are no observed option strings
to build dictionaries from. They pass through unencoded for now.

```{r}
#| label: pathway-d-check
survey |>
  select(any_of(c("q2d", "q3d", "q4d"))) |>
  summarise(across(everything(), \(x) sum(!is.na(x))))
```

## Single-select questions

Ordered categories become factors. Unobserved levels are filled in from the
questionnaire following the formatting pattern of the observed values;
`to_factor()` errors on any value not in its level set, so wording drift in a
future export surfaces immediately.

```{r}
#| label: factors
to_factor <- function(x, levels) {
  unknown <- setdiff(unique(x[!is.na(x)]), levels)
  if (length(unknown) > 0) {
    stop(
      "Value(s) outside the expected levels — add or fix:\n",
      paste0('  "', unknown, '"', collapse = "\n")
    )
  }
  factor(x, levels = levels)
}

survey <- survey |>
  mutate(
    hh_size = to_factor(hh_size, c(
      "1, just myself", "2", "3", "4", "5", "6", "7 or more",
      "Prefer not to say"
    )),
    hh_income = to_factor(hh_income, c(
      "Under $25,000", "$25,000-$49,999", "$50,000-$74,999",
      "$75,000-$99,999", "$100,000-$149,999", "$150,000-$174,999",
      "$175,000-$199,999", "$200,000-$249,999", "$250,000-$299,999",
      "$300,000 or more", "Prefer not to say"
    )),
    age = to_factor(age, c(
      "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older",
      "Prefer not to say"
    )),
    disability = to_factor(disability, c(
      "Yes", "No", "Not sure", "Prefer not to say"
    )),
    political = to_factor(political, c(
      "Very liberal", "Somewhat liberal", "Moderate/Middle of the road",
      "Somewhat conservative", "Very conservative", "Don't know",
      "Prefer not to say"
    ))
  )
```

## Pathway

Each respondent follows exactly one pathway through the survey, determined by
q1 (see the Survey Pathways table in the questionnaire). `derive_pathway()` in
`00_utils.R` assigns it with priority current > past > future — verified
against the data: a respondent eligible for both past and future was shown
only the past questions.

```{r}
#| label: pathway
survey <- derive_pathway(survey)

count(survey, pathway)
```

## Skip-logic checks

Each pathway question should be answered by exactly the respondents on its
pathway. Mismatches warn (rather than stop) — they signal an export or
routing problem to investigate, not necessarily a pipeline bug.

```{r}
#| label: skip-checks
skip_rules <- tribble(
  ~question, ~indicator,         ~expected_pathway,
  "q2a",     "q2a_unaffordable", "current",
  "q3a",     "q3a_other",        "current",
  "q4a",     "q4a_other",        "current",
  "q2b",     "q2b_job",          "future",
  "q3b",     "q3b_financial",    "future",
  "q2c",     "q2c_job",          "past",
  "q3c",     "q3c_backup",       "past"
)

pwalk(skip_rules, \(question, indicator, expected_pathway) {
  mismatch <- xor(!is.na(survey[[indicator]]), survey$pathway == expected_pathway)
  if (any(mismatch)) {
    cli::cli_warn(
      "{question}: {sum(mismatch)} respondent{?s} answered or skipped against
       skip logic (ID{?s} {.val {survey$respondent_id[mismatch]}})"
    )
  }
})
```

## Remaining columns

```{r}
#| label: other-cols
dupes <- survey$respondent_id[duplicated(survey$respondent_id)]
if (length(dupes) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes)}}")
}

survey <- survey |>
  mutate(
    # tz assumed UTC — the platform's actual export timezone is unconfirmed
    submitted_at = excel_datetime(submitted_at),
    zip = str_pad(zip, width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at)

bad_zip <- survey |> filter(!str_detect(zip, "^\\d{5}$") | is.na(zip))
if (nrow(bad_zip) > 0) {
  cli::cli_warn("{nrow(bad_zip)} row{?s} with missing or non-5-digit ZIP: {.val {bad_zip$zip}}")
}
```

## Validate and write

```{r}
#| label: check
glimpse(survey)
```

```{r}
#| label: write
write_rds(survey, here("data/processed/survey_clean.rds"))
write_csv(survey, here("data/processed/survey_clean.csv"))
```

## Canvassing data

The canvassing form is a doorstep funnel — canvasser logistics, a short
screener, then (for those who agree to continue) the same question battery as
the digital survey, with identical question and option text, so the survey's
column map and dictionaries are reused. The canvasser-only doorstep challenge
question uses its own short labels. There are no demographic questions and no
respondent ZIP; the only geography is the *canvasser's* ZIP.

```{r}
#| label: canvassing-read
canvassing_files <- list.files(
  here("data/raw/canvassing"),
  pattern = "\\.xlsx$", full.names = TRUE
) |>
  str_subset("~\\$", negate = TRUE)

canvassing_raw <- canvassing_files |>
  map(read_excel, col_types = "text") |>
  bind_rows() |>
  mutate(across(everything(), \(x) na_if(str_trim(x), "")))

dim(canvassing_raw)
```

```{r}
#| label: canvassing-rename
battery <- col_map[c(
  "q1", "q2a", "q3a", "q4a", "q2b", "q3b", "q2c", "q3c",
  "q2d", "q3d", "q4d", "q6"
)]

col_map_canv <- c(
  submitted_at        = "Time",
  canvasser_zip       = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Zip)",
  canvass_mode        = "For Canvasser: How are you conducting this conversation?",
  responded           = "For Canvasser: Did the person respond and agree to answer questions?  If no, share the information card or the careconversations.com website and submit.",
  care_connection     = "Have you provided, received care yourself, expected to in the future or do you know someone who is? This could be giving or receiving care for an aging family member, a friend, a child, or someone with a disability no matter their age. It also includes you.",
  doorstep_challenges = "What challenges, if any, have you or others faced — or do you expect to face — when it comes to giving, finding, or getting care? (For canvasser: Check all that apply. If selecting 'No Challenges,' leave everything else unchecked.)",
  continue_full       = "We're talking with people across the country about their experiences with care — what's working and what families need.  Can you take 5-minutes to share more?",
  battery,
  q4c                 = "What or who helped you most with care? (Choose your top 3)",
  respondent_id       = "Unique ID"
)

canvassing <- rename_validated(canvassing_raw, col_map_canv)
```

The doorstep challenge question gets its own dictionary; the shared battery
reuses the survey dictionaries. q4c (like q2d/q3d/q4d) has no responses in the
test data yet, so it passes through unencoded.

```{r}
#| label: canvassing-encode
doorstep_options <- c(
  cost                = "Cost",
  time                = "Time",
  location            = "Location",
  right_fit           = "Right fit",
  lack_support        = "Lack of support",
  no_control          = "No control",
  worker_availability = "Worker availability",
  barriers_access     = "Barriers to access",
  no_challenges       = "No challenges"
)

canv_multiselect <- c(
  list(doorstep_challenges = doorstep_options),
  multiselect_options[c("q1", "q2a", "q3a", "q4a", "q2b", "q3b", "q2c", "q3c", "q6")]
)

canvassing <- reduce(
  names(canv_multiselect),
  \(data, col) encode_multiselect(data, col, canv_multiselect[[col]]),
  .init = canvassing
) |>
  derive_pathway()
```

```{r}
#| label: canvassing-clean
canvassing <- canvassing |>
  mutate(
    canvass_mode    = to_factor(canvass_mode, c("Door-to-door", "Phone", "Other In Person")),
    responded       = to_factor(responded, c("Yes - continue with questions", "No - not available")),
    care_connection = to_factor(care_connection, c("Yes", "No")),
    continue_full   = to_factor(continue_full, c("Yes - continuing", "No - not continuing")),
    submitted_at    = excel_datetime(submitted_at),
    canvasser_zip   = str_pad(canvasser_zip, width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at)

dupes_canv <- canvassing$respondent_id[duplicated(canvassing$respondent_id)]
if (length(dupes_canv) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes_canv)}}")
}

# the battery should be answered by exactly those who agreed to continue
agreed <- !is.na(canvassing$continue_full) & canvassing$continue_full == "Yes - continuing"
battery_answered <- !is.na(canvassing$q1_child_now)
if (any(xor(agreed, battery_answered))) {
  cli::cli_warn("Battery answered/skipped against the doorstep funnel for
                 {sum(xor(agreed, battery_answered))} respondent{?s}")
}

count(canvassing, responded, care_connection, continue_full, pathway)
```

```{r}
#| label: canvassing-write
write_rds(canvassing, here("data/processed/canvassing_clean.rds"))
write_csv(canvassing, here("data/processed/canvassing_clean.csv"))
```

## Data dictionary

The column maps and option dictionaries above already document the schema —
this writes them out as a reviewable CSV: one row per indicator column (or
factor level), with the source question and option text.

```{r}
#| label: dictionary
question_text <- c(col_map, col_map_canv[setdiff(names(col_map_canv), names(col_map))])
all_multiselect <- c(multiselect_options, list(doorstep_challenges = doorstep_options))

dict_options <- all_multiselect |>
  imap(\(opts, q) tibble(
    column   = str_c(q, "_", names(opts)),
    question = unname(question_text[q]),
    option   = unname(opts)
  )) |>
  list_rbind()

dict_factors <- list(survey, canvassing) |>
  map(\(data) {
    data |>
      select(where(is.factor)) |>
      as.list() |>
      imap(\(v, nm) tibble(
        column   = nm,
        question = unname(question_text[nm]),
        option   = levels(v)
      )) |>
      list_rbind()
  }) |>
  list_rbind() |>
  distinct()

dictionary <- bind_rows(dict_options, dict_factors) |>
  mutate(question = if_else(
    column == "pathway",
    "Derived from q1 (priority: current > past > future > observer > none)",
    question
  ))

write_csv(dictionary, here("data/processed/data-dictionary.csv"))
dictionary
```

## Long view

The wide file is canonical (one row per respondent, the unit of analysis).
For option-level summaries and plots, `pivot_selections()` in
`scripts/00_utils.R` reshapes it to one row per respondent × option:

```{r}
#| label: long-view
survey |>
  pivot_selections() |>
  summarise(rate = mean(selected, na.rm = TRUE), .by = c(question, option)) |>
  arrange(question, desc(rate))
```
