---
title: "Clean test data"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
    code-summary: "Show code"
execute:
  warning: true
  message: false
---

Cleans the raw survey and canvassing exports into analysis-ready files in
`data/processed/`, plus a generated data dictionary.

`data/raw/Raw data backups/` is a local mirror of the Box delivery folder,
kept exactly as Box structures it — the filenames already encode source
(`surv`/`canv`), collection dates, form version (`f0`, `f1`, ...), and
complete vs partial, and the accompanying `_ DATA LOG _.xlsx` is the delivery
manifest keyed to those filenames. Files are selected by name pattern; nothing
is re-filed by hand. The `f0` files are the pre-launch test delivery (per the
data log) and are excluded; survey cleaning covers real deliveries (`f1`+),
with complete and partial responses combined and distinguished by a
`response_status` column.

The main transformation is one-hot encoding the multi-select questions: the raw
export stores multiple selections as a single pipe-delimited string, and we
break each question into one logical column per response option. `TRUE` means
the option was selected, `FALSE` means the respondent saw the question but did
not select it, and `NA` means the respondent never saw the question (skip
logic) or skipped it. Selection rates are just `mean(x, na.rm = TRUE)`.

Response option text comes from the observed data, not the survey PDF — the
live form's wording differs from the June 15 questionnaire document (e.g. the
PDF's "Caring for a child/children" appears in the data as "I care for a child
or children now"). Options not yet observed in the data are therefore not in
the dictionaries below; the encoding stops with an error listing any new value
so it can be added deliberately.

**Warnings in this render are expected and monitored, not failures.** They
are the drift alarms: known skip-logic exceptions (the paid-provider
dual-battery routing), ZIP+4 entries, and the one canvassing funnel
violation — all documented in `SURVEY.md`. A *new* warning appearing here is
the signal to investigate.

```{r}
#| label: setup
library(tidyverse)
library(readxl)
library(here)

source(here("scripts/00_utils.R"))
```

## Read raw survey data

```{r}
#| label: read
raw_dir <- here("data/raw/Raw data backups")

survey_files <- list.files(
  raw_dir,
  # deliveries from 2026-08 on carry a "nopii_" prefix (PII-skimmed upstream)
  pattern = "^(nopii_)?surv_export_.*_closed.*\\.xlsx$", full.names = TRUE
) |>
  str_subset("~\\$", negate = TRUE) |> # Excel lock files, present while a workbook is open
  str_subset("_f0_", negate = TRUE)    # f0 = pre-launch test delivery per the data log

# The anchored pattern above is deliberate — but a future prefix change would
# make it silently skip deliveries. Any closed survey export it didn't claim
# (f0 aside) is treated as naming drift and stops the render.
unclaimed <- list.files(raw_dir, pattern = "surv_export_.*_closed.*\\.xlsx$", full.names = TRUE) |>
  str_subset("~\\$", negate = TRUE) |>
  str_subset("_f0_", negate = TRUE) |>
  setdiff(survey_files)
if (length(unclaimed) > 0) {
  cli::cli_abort(c(
    "{length(unclaimed)} closed survey export{?s} not matched by the filename pattern — naming drift? Update the pattern deliberately:",
    set_names(basename(unclaimed), "x")
  ))
}

basename(survey_files)
```

## Rename columns

Column names in the export are the full question text. We map them to short
IDs keyed to the questionnaire (q1, q2a, ..., q6). The mapping is validated in
both directions so a new export with added, dropped, or reworded columns fails
loudly instead of silently misaligning; columns known to vary across form
versions (the f0-only address fields, q4c which f0 lacked) are marked
optional. Delivery metadata parsed from each filename travels with every row.

The open-text questions (q5a/q5c/q5d, q7) come through a separate export and
will need entries here when they join the closed-text pipeline.

```{r}
#| label: rename
col_map <- c(
  submitted_at  = "Time",
  q1            = "We would like to know more about your connection to care, if any. Choose all that apply to you.",
  q2a           = "What challenges, if any, have you experienced with providing, finding, or receiving care? (Choose all that apply)",
  q3a           = "What care support would make a difference for you or people around you? (Choose all that apply)",
  q4a           = "What or who helps you most with care? (Choose your top 3)",
  q2b           = "When you think about the care you, a family member, or friend might need 5 years from now, what challenges, if any, concern you most? (Choose all that apply)",
  q3b           = "What types of care support and services do you think will matter most for you or people you know? (Choose all that apply)",
  q2c           = "When you provided or received care what challenges, if any, did you experience? (Choose all that apply)",
  q3c           = "What care support would have made a difference for you or people around you? (Choose all that apply)",
  q2d           = "What care challenges, if any, do the people you know experience? (Choose all that apply.)",
  q3d           = "What care support would make a difference for the people you know? (Choose all that apply)",
  q4d           = "What or who helps most with their care? (Choose your top 3)",
  q4c           = "What or who helped you most with care? (Choose your top 3)",
  q6            = "What would ideal care support and services look like for you? (Choose up to three)",
  address       = "Zipcode (Address)",
  address2      = "Zipcode (Address2)",
  city          = "Zipcode (City)",
  state         = "Zipcode (State)",
  zip           = "Zipcode (Zip)",
  hh_size       = "How many people live in your household? (Include yourself and anyone you share income and living expenses with, including kids)",
  hh_income     = "What is your household's approximate annual income? (Add up income from everyone living with you, related or not.)",
  age           = "What is your age?",
  gender        = "What is your gender? (Choose all that apply)",
  race          = "What is your race or ethnicity? (Choose all that apply)",
  disability    = "I identify as a person with a disability or chronic health condition.",
  political     = "How would you describe your political views?",
  respondent_id = "Unique ID"
)

form_varying <- c("address", "address2", "city", "state", "q4c")

read_delivery <- function(path) {
  read_excel(path, col_types = "text") |>
    mutate(across(everything(), \(x) na_if(str_trim(x), ""))) |> # whitespace-only cells -> NA
    rename_validated(col_map, optional = form_varying) |>
    mutate(delivery = basename(path))
}

survey <- survey_files |>
  map(read_delivery) |>
  bind_rows() |>
  mutate(
    form_version    = str_extract(delivery, "f\\d+"),
    response_status = str_extract(delivery, "complete|partial")
  )

count(survey, delivery, form_version, response_status)
```

## PII guard

Data shared with Urban must contain no PII. ZIP is collected by design, but
the export includes street-address fields that should always be empty — stop
immediately if they ever are not.

```{r}
#| label: pii-guard
pii_cols <- intersect(c("address", "address2"), names(survey))
if (length(pii_cols) > 0) {
  if (any(!is.na(as.matrix(survey[pii_cols])))) {
    stop("Address fields are not empty — possible PII in the raw file. Do not proceed; flag to the team.")
  }
  survey <- select(survey, -all_of(pii_cols))
}
```

## One-hot encode multi-select questions

Selections arrive pipe-delimited in survey order, e.g.
`"Care is unaffordable|Other: my own words"`. Each question becomes one
logical column per option (`q2a_unaffordable`, ...), with every option the
respondent did not select set to `FALSE` and rows where the question was
unanswered set to `NA` across all of its option columns.

"Other" responses are open form in the real data, and Formstack exports them
three ways: `"Other: <text>"`, a bare `"Other"`/`"true"` (the checkbox with no
text), or occasionally the free text alone with no prefix. Any token that
isn't a known structured option sets `{question}_other = TRUE` and lands
verbatim in `{question}_other_text` — that text is respondent voice and feeds
the qualitative/NLP work. As a drift alarm, a free-text value repeating across
3+ respondents triggers a warning, since repetition suggests a new
*structured* option missing from the dictionary rather than genuine free text.

A dictionary entry may list several accepted wordings for the same option: the
August 19, 2026 form revision reworded a handful of options without changing
their meaning, and the research team combines old and new responses, so any
listed variant sets the same indicator column.

`encode_multiselect()` lives in `scripts/00_utils.R`. A quick demonstration
on a toy question: `"c | e"` gets `TRUE` for c and e and `FALSE` for
everything else; an "Other" with text sets the other flag and captures the
text; an unanswered row is `NA` throughout; and option `a` accepts two
wordings (an alias, as used for the Aug 19 revision) — either one sets
`q_demo_a`.

```{r}
#| label: encode-demo
tibble(q_demo = c("c | e", "a-reworded|Other: my own answer", "a", NA)) |>
  encode_multiselect(
    "q_demo",
    list(a = c("a", "a-reworded"), b = "b", c = "c", d = "d", e = "e", f = "f")
  )
```

```{r}
#| label: dictionaries
multiselect_options <- list(
  q1 = c(
    child_now      = "I care for a child or children now",
    aging_now      = "I care for an aging person now",
    disability_now = "I care for someone with a disability or chronic health condition now",
    paid_provider  = "I am a paid care provider. I offer support and services for kids, aging adults, and/or people with disabilities",
    need_care_now  = "I need care but am not receiving it",
    receives_care  = "I receive or have received care for a disability or chronic condition",
    cared_past     = "I have cared for an aging adult, child, or someone with a disability or chronic health condition in the past",
    expect_future  = "I am not providing or receiving care right now, but I expect to in the next 5 years",
    observer       = "None of these apply to me right now, but I know someone who receives or needs care",
    none           = "None of these apply to me"
  ),
  q2a = list(
    unaffordable = "Care is unaffordable",
    finding      = c(
      "It is hard to find care that's available, nearby, or right for my family or me",
      # Aug 19 2026 revision — seeded from the questionnaire PDF, not yet observed in an export
      "It's hard to find care that's available, nearby, or right for me or my family"
    ),
    time_flex    = "I don't have enough time or flexibility",
    help         = "I don't have enough help or support",
    figure_out   = "It is hard to figure out what help exists or how to get it",
    navigate     = "It is hard to navigate existing care programs",
    no_say       = "I don't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3a = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    everyday_support   = "Everyday support like meals and transportation",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4a = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, child care, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid for using federal, state, or county dollars",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I do not receive help"
  ),
  q2b = c(
    unaffordable = "Care that's unaffordable",
    finding      = "Difficulty with finding care that's available, nearby, or right for me or my family",
    time_flex    = "Not having time or flexibility",
    help         = "Not enough help or support with care",
    figure_out   = "Difficulty knowing what help exists or how to get it",
    navigate     = "Difficulty navigating existing care programs",
    no_say       = "Not having a say in how care works for my family or me",
    job          = "Care could affect my job or career"
  ),
  q3b = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q2c = c(
    unaffordable = "Care was unaffordable",
    finding      = "It was hard to find care that was available, nearby, or right for my family and me",
    time_flex    = "I didn't have enough time or flexibility",
    help         = "I didn't have enough help or support",
    figure_out   = "It was hard to figure out what help existed or how to get it",
    navigate     = "It was hard to navigate existing care programs",
    no_say       = "I didn't have a say in how care works for me or my family",
    job          = "Care affected my job or career"
  ),
  q3c = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4c = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  q2d = c(
    unaffordable = "Unaffordable care",
    finding      = "It was hard to find care that's available, nearby, or right for them",
    time_flex    = "They don't have enough time or flexibility",
    help         = "They are doing this without enough help or support",
    figure_out   = "It is hard for them to figure out what help exists or how to get it",
    navigate     = "It is hard for them to navigate existing care programs",
    no_say       = "They don't have a say in how care works for them",
    job          = "Care has affected their jobs or careers"
  ),
  q3d = c(
    affordable_quality = "Affordable, high quality care for an aging adult, a child, someone with a disability, or myself",
    backup             = "Backup care for emergencies or unexpected needs",
    extended_childcare = "Extended child care options such as before 9:00AM, after 5:00PM, or during the summer",
    financial          = "Financial support for care expenses",
    navigation_help    = "Help with finding, understanding, and accessing available care options",
    schedule_control   = "More control over my schedule at work",
    paid_time_off      = "Paid time off for care",
    respite            = "Respite care"
  ),
  q4d = c(
    family_friends       = "Family, friends, or neighbors",
    health_professionals = "Health care professionals",
    paid_provider        = "Paid care provider for home care or personal care, daycare, childcare, or pre-k",
    community_religious  = "Community organizations or religious groups",
    government           = "Government support, like paid leave or other care services paid with federal, state, or county money",
    # in the instrument (per the Aug 19 2026 PDF, wording as in q4c) but unobserved so far
    employer             = "Employer support for things like leave, flexibility, benefits",
    support_groups       = "Support groups, online, or in-person",
    technology           = "Technology or online tools",
    no_help              = "I did not receive help"
  ),
  # Aug 19 2026 variants are seeded from the questionnaire PDF with the
  # formatting pattern of observed exports (plain hyphens, not the PDF's
  # em-dashes) — confirm against the first export that carries them
  q6 = list(
    there_when_needed = "There when you need it - at every stage of life, no matter your job or family situation",
    easy_to_use       = c(
      "Easy to find and use - without complicated steps or barriers",
      "Easy to find and use - without confusing steps or barriers"          # Aug 19 2026
    ),
    affordable        = "Affordable - without causing personal financial strain",
    in_community      = "Available in your community - not far away or out of reach",
    flexible          = "Flexible - fitting real schedules, needs, and changing circumstances",
    reliable          = "Reliable - something you can count on over time",
    clear             = "Clear and transparent - you know what's available and how to get it",
    good_jobs         = "Good jobs for care providers - with fair pay and stability",
    healthy_dev       = c(
      "Support healthy development, well-being, and independence",
      "Supports healthy development, well-being, and independence"          # Aug 19 2026
    ),
    lived_experience  = c(
      "Shaped by people with experience giving, receiving, or needing care",
      "Shaped by people with lived care experience"                         # Aug 19 2026
    ),
    secure            = "Keep people who rely on care feeling secure, healthy, and thriving"
  ),
  gender = c(
    woman             = "Woman",
    man               = "Man",
    nonbinary         = "Non-binary",
    transgender       = "Transgender",
    prefer_not_to_say = "Prefer not to say"
  ),
  race = c(
    aian              = "American Indian or Alaska Native",
    asian             = "Asian or Asian American",
    black             = "Black or African American",
    hispanic          = "Hispanic, Latino, or Latina",
    mena              = "Middle Eastern or North African",
    nhpi              = "Native Hawaiian or Pacific Islander",
    white             = "White",
    multiracial       = "Multiracial",
    prefer_not_to_say = "Prefer not to say"
  )
)

survey <- reduce(
  names(multiselect_options),
  \(data, col) encode_multiselect(data, col, multiselect_options[[col]]),
  .init = survey
)
```


## Single-select questions

Ordered categories become factors. Unobserved levels are filled in from the
questionnaire following the formatting pattern of the observed values;
`to_factor()` errors on any value not in its level set, so wording drift in a
future export surfaces immediately.

```{r}
#| label: factors
to_factor <- function(x, levels) {
  unknown <- setdiff(unique(x[!is.na(x)]), levels)
  if (length(unknown) > 0) {
    stop(
      "Value(s) outside the expected levels — add or fix:\n",
      paste0('  "', unknown, '"', collapse = "\n")
    )
  }
  factor(x, levels = levels)
}

survey <- survey |>
  mutate(
    # f1 exports numeric sizes as "2.0", "3.0", ...
    hh_size = to_factor(str_remove(hh_size, "\\.0$"), c(
      "1, just myself", "2", "3", "4", "5", "6", "7 or more",
      "Prefer not to say"
    )),
    hh_income = to_factor(hh_income, c(
      "Under $25,000", "$25,000-$49,999", "$50,000-$74,999",
      "$75,000-$99,999", "$100,000-$149,999", "$150,000-$174,999",
      "$175,000-$199,999", "$200,000-$249,999", "$250,000-$299,999",
      "$300,000 or more", "Prefer not to say"
    )),
    age = to_factor(age, c(
      "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older",
      "Prefer not to say"
    )),
    disability = to_factor(disability, c(
      "Yes", "No", "Not sure", "Prefer not to say"
    )),
    political = to_factor(political, c(
      "Very liberal", "Somewhat liberal", "Moderate/Middle of the road",
      "Somewhat conservative", "Very conservative", "Don't know",
      "Prefer not to say"
    ))
  )
```

## Pathway

Each respondent follows exactly one pathway through the survey, determined by
q1 (see the Survey Pathways table in the questionnaire). `derive_pathway()` in
`00_utils.R` assigns it with priority current > past > future — verified
against the data: a respondent eligible for both past and future was shown
only the past questions.

```{r}
#| label: pathway
survey <- derive_pathway(survey)

count(survey, pathway)
```

## Skip-logic checks

Each pathway question should be answered by exactly the respondents on its
pathway. For complete responses the check runs in both directions; partial
responses may legitimately stop mid-survey, so they are only flagged when they
*answered* a question outside their pathway. Mismatches warn (rather than
stop) — they signal an export or routing problem to investigate, not
necessarily a pipeline bug.

```{r}
#| label: skip-checks
skip_rules <- tribble(
  ~question, ~indicator,             ~expected_pathway,
  "q2a",     "q2a_unaffordable",     "current",
  "q3a",     "q3a_backup",           "current",
  "q4a",     "q4a_family_friends",   "current",
  "q2b",     "q2b_job",              "future",
  "q3b",     "q3b_financial",        "future",
  "q2c",     "q2c_job",              "past",
  "q3c",     "q3c_backup",           "past",
  "q4c",     "q4c_family_friends",   "past",
  "q2d",     "q2d_unaffordable",     "observer",
  "q3d",     "q3d_backup",           "observer",
  "q4d",     "q4d_family_friends",   "observer"
)

# q6 is shown to every pathway, so it has no row in skip_rules — instead a
# complete response that skipped it entirely signals routing/export drift
q6_skipped <- survey$response_status == "complete" & is.na(survey$q6_affordable)
if (any(q6_skipped)) {
  cli::cli_warn("q6: {sum(q6_skipped)} complete response{?s} with q6 entirely missing")
}

pwalk(skip_rules, \(question, indicator, expected_pathway) {
  answered <- !is.na(survey[[indicator]])
  eligible <- survey$pathway %in% expected_pathway
  mismatch <- if_else(
    survey$response_status == "complete",
    xor(answered, eligible), # completes must answer exactly their pathway
    answered & !eligible     # partials may stop early, but not cross paths
  )
  n_bad <- sum(mismatch, na.rm = TRUE)
  if (n_bad > 0) {
    cli::cli_warn(
      "{question}: {n_bad} respondent{?s} answered or skipped against skip
       logic (e.g. ID{?s} {.val {head(survey$respondent_id[which(mismatch)], 5)}})"
    )
  }
})
```

## Remaining columns

```{r}
#| label: other-cols
survey <- survey |>
  mutate(
    respondent_id = normalize_id(respondent_id),
    # tz assumed UTC — the platform's actual export timezone is unconfirmed
    submitted_at = excel_datetime(submitted_at),
    # same "91401.0" float artifact as hh_size; was invisible while warnings were silenced
    zip = str_pad(str_remove(zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

# after normalization so ID collisions across export formats would surface
dupes <- survey$respondent_id[duplicated(survey$respondent_id)]
if (length(dupes) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes)}}")
}

# missing ZIPs are expected for partials that stopped before demographics
bad_zip <- survey |> filter(!is.na(zip), !str_detect(zip, "^\\d{5}$"))
if (nrow(bad_zip) > 0) {
  cli::cli_warn("{nrow(bad_zip)} row{?s} with non-5-digit ZIP: {.val {unique(bad_zip$zip)}}")
}
count(survey, response_status, zip_missing = is.na(zip))
```

## Validate and write

```{r}
#| label: check
glimpse(survey)
```

```{r}
#| label: write
write_rds(survey, here("data/processed/survey_clean.rds"))
write_csv(survey, here("data/processed/survey_clean.csv"))
```

## Canvassing data

The canvassing form is a doorstep funnel — canvasser logistics, a short
screener, then (for those who agree to continue) the same question battery as
the digital survey, with identical question and option text, so the survey's
column map and dictionaries are reused. The canvasser-only doorstep challenge
question uses its own short labels. There are no demographic questions and no
respondent ZIP; the only geography is the *canvasser's* ZIP.

Like the survey, the `f0` pre-launch test is excluded and cleaning covers real
deliveries (`f1`+, complete and partial combined). The `*_opentext` exports
(here and for the survey) are set aside for the text-analysis pipeline and not
read by this script.

```{r}
#| label: canvassing-read
canvassing_files <- list.files(
  raw_dir,
  pattern = "^(nopii_)?canv_export_.*_closed.*\\.xlsx$", full.names = TRUE
) |>
  str_subset("~\\$", negate = TRUE) |>
  str_subset("_f0_", negate = TRUE)

# same naming-drift guard as the survey: a closed canvassing export the
# pattern didn't claim stops the render instead of being silently skipped
unclaimed_canv <- list.files(raw_dir, pattern = "canv_export_.*_closed.*\\.xlsx$", full.names = TRUE) |>
  str_subset("~\\$", negate = TRUE) |>
  str_subset("_f0_", negate = TRUE) |>
  setdiff(canvassing_files)
if (length(unclaimed_canv) > 0) {
  cli::cli_abort(c(
    "{length(unclaimed_canv)} closed canvassing export{?s} not matched by the filename pattern — naming drift? Update the pattern deliberately:",
    set_names(basename(unclaimed_canv), "x")
  ))
}

basename(canvassing_files)
```

```{r}
#| label: canvassing-rename
battery <- col_map[c(
  "q1", "q2a", "q3a", "q4a", "q2b", "q3b", "q2c", "q3c",
  "q2d", "q3d", "q4d", "q6"
)]

col_map_canv <- c(
  submitted_at        = "Time",
  canvasser_zip       = "For Canvasser: Enter the ZIP code of the place you are canvassing  (Zip)",
  canvass_mode        = "For Canvasser: How are you conducting this conversation?",
  responded           = "For Canvasser: Did the person respond and agree to answer questions?  If no, share the information card or the careconversations.com website and submit.",
  care_connection     = "Have you provided, received care yourself, expected to in the future or do you know someone who is? This could be giving or receiving care for an aging family member, a friend, a child, or someone with a disability no matter their age. It also includes you.",
  doorstep             = "What challenges, if any, have you or others faced — or do you expect to face — when it comes to giving, finding, or getting care? (For canvasser: Check all that apply. If selecting 'No Challenges,' leave everything else unchecked.)",
  continue_full       = "We're talking with people across the country about their experiences with care — what's working and what families need.  Can you take 5-minutes to share more?",
  battery,
  q4c                 = "What or who helped you most with care? (Choose your top 3)",
  respondent_id       = "Unique ID"
)

canvassing <- canvassing_files |>
  map(\(path) {
    read_excel(path, col_types = "text") |>
      mutate(across(everything(), \(x) na_if(str_trim(x), ""))) |>
      rename_validated(col_map_canv) |>
      mutate(delivery = basename(path))
  }) |>
  bind_rows() |>
  mutate(
    form_version    = str_extract(delivery, "f\\d+"),
    response_status = str_extract(delivery, "complete|partial")
  )

count(canvassing, delivery, form_version, response_status)
```

The doorstep challenge question gets its own dictionary; the shared battery
reuses the survey dictionaries for every pathway question.

```{r}
#| label: canvassing-encode
doorstep_options <- c(
  cost                = "Cost",
  time                = "Time",
  location            = "Location",
  right_fit           = "Right fit",
  lack_support        = "Lack of support",
  no_control          = "No control",
  worker_availability = "Worker availability",
  barriers_access     = "Barriers to access",
  no_challenges       = "No challenges",
  # added to the form between f0 and f1 (surfaced by the repeat-drift alarm)
  availability        = "Availability",
  figure_out          = "Hard to figure out"
)

canv_multiselect <- c(
  list(doorstep = doorstep_options),
  multiselect_options[c(
    "q1", "q2a", "q3a", "q4a", "q2b", "q3b",
    "q2c", "q3c", "q4c", "q2d", "q3d", "q4d", "q6"
  )]
)

canvassing <- reduce(
  names(canv_multiselect),
  \(data, col) encode_multiselect(data, col, canv_multiselect[[col]]),
  .init = canvassing
) |>
  derive_pathway()
```

```{r}
#| label: canvassing-clean
canvassing <- canvassing |>
  mutate(
    respondent_id   = normalize_id(respondent_id),
    canvass_mode    = to_factor(canvass_mode, c("Door-to-door", "Phone", "Other In Person")),
    responded       = to_factor(responded, c("Yes - continue with questions", "No - not available")),
    care_connection = to_factor(care_connection, c("Yes", "No")),
    continue_full   = to_factor(continue_full, c("Yes - continuing", "No - not continuing")),
    submitted_at    = excel_datetime(submitted_at),
    # same float artifact as the survey ZIPs
    canvasser_zip   = str_pad(str_remove(canvasser_zip, "\\.0$"), width = 5, pad = "0")
  ) |>
  relocate(respondent_id, submitted_at, form_version, response_status, delivery)

dupes_canv <- canvassing$respondent_id[duplicated(canvassing$respondent_id)]
if (length(dupes_canv) > 0) {
  cli::cli_abort("Duplicated respondent ID{?s}: {.val {unique(dupes_canv)}}")
}

# the battery should be answered by exactly those who agreed to continue;
# partials may legitimately stop before the battery, but not answer it
# without having agreed
agreed <- !is.na(canvassing$continue_full) & canvassing$continue_full == "Yes - continuing"
battery_answered <- !is.na(canvassing$q1_child_now)
funnel_mismatch <- if_else(
  canvassing$response_status == "complete",
  xor(agreed, battery_answered),
  battery_answered & !agreed
)
if (any(funnel_mismatch)) {
  cli::cli_warn("Battery answered/skipped against the doorstep funnel for
                 {sum(funnel_mismatch)} respondent{?s}")
}

count(canvassing, response_status, responded, care_connection, continue_full, pathway)
```

```{r}
#| label: canvassing-write
write_rds(canvassing, here("data/processed/canvassing_clean.rds"))
write_csv(canvassing, here("data/processed/canvassing_clean.csv"))
```

## Data dictionary

The column maps and option dictionaries above already document the schema —
this writes them out as a reviewable CSV: one row per indicator column (or
factor level), with the source question and option text.

```{r}
#| label: dictionary
question_text <- c(col_map, col_map_canv[setdiff(names(col_map_canv), names(col_map))])
all_multiselect <- c(multiselect_options, list(doorstep = doorstep_options))

dict_options <- all_multiselect |>
  imap(\(opts, q) tibble(
    column   = str_c(q, "_", c(names(opts), "other", "other_text")),
    question = unname(question_text[q]),
    # aliased options (form revisions) list every accepted wording
    option   = c(unname(map_chr(opts, str_flatten, collapse = " [or] ")),
                 "Other (any response not in the structured options)",
                 "Free text captured from other responses")
  )) |>
  list_rbind()

dict_factors <- list(survey, canvassing) |>
  map(\(data) {
    data |>
      select(where(is.factor)) |>
      as.list() |>
      imap(\(v, nm) tibble(
        column   = nm,
        question = unname(question_text[nm]),
        option   = levels(v)
      )) |>
      list_rbind()
  }) |>
  list_rbind() |>
  distinct()

dictionary <- bind_rows(dict_options, dict_factors) |>
  mutate(question = if_else(
    column == "pathway",
    "Derived from q1 (priority: current > past > future > observer > none)",
    question
  ))

write_csv(dictionary, here("data/processed/data-dictionary.csv"))
dictionary
```

## Long view

The wide file is canonical (one row per respondent, the unit of analysis).
For option-level summaries and plots, `pivot_selections()` in
`scripts/00_utils.R` reshapes it to one row per respondent × option:

```{r}
#| label: long-view
survey |>
  pivot_selections() |>
  summarise(rate = mean(selected, na.rm = TRUE), .by = c(question, option)) |>
  arrange(question, desc(rate))
```
