---
title: "Exploring the 1MCC test data"
format:
  html:
    embed-resources: true
    toc: true
    code-fold: true
    code-summary: "Show code"
execute:
  warning: false
  message: false
fig-width: 7
fig-height: 4
---

A first look at what the 1M Care Conversations data contains and the kinds of
analysis it will support. The survey numbers are now **real**: the first live
delivery covers June 4 – July 10, 2026, with 327 complete and 158 partial
responses. Canvassing remains the pre-launch **test** export (12 contacts), so
survey and canvassing are shown separately and never pooled; canvassing charts
illustrate the machinery, not findings. The same code runs unchanged as new
deliveries land in `data/raw/`.

Data comes from `scripts/01_clean-data.qmd`, which documents the cleaning; the
generated `data/processed/data-dictionary.csv` maps every column to its
question and option text.

```{r}
#| label: setup
library(tidyverse)
library(here)
library(urbnthemes)
library(scales)

set_urbn_defaults(style = "print")
source(here("scripts/00_utils.R"))

survey     <- read_rds(here("data/processed/survey_clean.rds"))
canvassing <- read_rds(here("data/processed/canvassing_clean.rds"))
dictionary <- read_csv(here("data/processed/data-dictionary.csv"), show_col_types = FALSE)

# long view of every multi-select response, both sources, with option text
responses <- bind_rows(
  survey     = pivot_selections(survey),
  canvassing = pivot_selections(canvassing),
  .id = "source"
) |>
  mutate(indicator = str_c(question, "_", option)) |>
  left_join(
    dictionary |> select(indicator = column, option_text = option),
    by = "indicator"
  ) |>
  mutate(option_text = str_wrap(coalesce(option_text, option), 40))
```

## The two data sources

The digital survey and the canvassing form share the same core question
battery (connection to care, challenges, supports, ideal care), so responses
can be analyzed jointly. They differ around the edges: the survey collects
demographics and respondent ZIP; canvassing instead records the doorstep
funnel and the canvasser's ZIP.

```{r}
#| label: overview
bind_rows(
  survey     = survey |>
    summarise(respondents = n(),
              first = min(as_date(submitted_at)), last = max(as_date(submitted_at)),
              .by = response_status),
  canvassing = canvassing |>
    mutate(response_status = "test") |>
    summarise(respondents = n(),
              first = min(as_date(submitted_at)), last = max(as_date(submitted_at)),
              .by = response_status),
  .id = "source"
) |>
  knitr::kable()
```

### The canvassing funnel

Canvassing is a contact funnel: not everyone approached answers the door, and
not everyone who answers agrees to the full questions. Analysis of canvassing
responses needs the right denominator for each stage.

```{r}
#| label: funnel
funnel <- tibble(
  stage = c(
    "Approached",
    "Answered and engaged",
    "Reports a connection to care",
    "Agreed to full questions"
  ),
  n = c(
    nrow(canvassing),
    sum(canvassing$responded == "Yes - continue with questions", na.rm = TRUE),
    sum(canvassing$care_connection == "Yes", na.rm = TRUE),
    sum(!is.na(canvassing$pathway))
  )
) |>
  mutate(stage = fct_rev(fct_inorder(stage)))

ggplot(funnel, aes(n, stage)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.5) +
  scale_x_continuous(expand = expansion(mult = c(0, .1))) +
  labs(
    title = "Canvassing doorstep funnel",
    subtitle = "Test data, June 2-8 2026",
    x = "Contacts", y = NULL
  )
```

## Who is answering: care pathways

Each respondent is routed down one pathway by their connection to care —
current caregivers/recipients get the fullest question set. Pathway
composition determines which questions have data and is the first thing to
watch as real deliveries arrive.

```{r}
#| label: pathways
survey |>
  filter(!is.na(pathway)) |>
  count(pathway, response_status) |>
  ggplot(aes(pathway, n, fill = response_status)) +
  geom_col() +
  geom_text(
    aes(label = after_stat(y), group = pathway),
    stat = "summary", fun = sum, vjust = -0.5
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, .1))) +
  labs(
    title = "Survey respondents by care pathway",
    subtitle = "Real survey data; 15 partials abandoned before q1 excluded",
    x = NULL, y = "Respondents", fill = NULL
  )
```

## What people report

The workhorse summary for every multi-select question is the selection rate:
the share of respondents *shown the question* who picked each option. With
real volume these become the headline descriptive statistics, with confidence
intervals and demographic/geographic breakdowns.

### Challenges among current caregivers and recipients (q2a)

```{r}
#| label: challenges
responses |>
  filter(source == "survey", question == "q2a", option != "other_text", !is.na(selected)) |>
  summarise(n_selected = sum(selected), n_shown = n(), .by = option_text) |>
  mutate(rate = n_selected / n_shown) |>
  ggplot(aes(rate, fct_reorder(option_text, rate))) +
  geom_col() +
  geom_text(aes(label = str_glue("{n_selected} of {n_shown}")), hjust = -0.2) +
  scale_x_continuous(labels = label_percent(), expand = expansion(mult = c(0, .15))) +
  labs(
    title = "Care challenges reported on the current pathway",
    subtitle = "Share of survey respondents shown the question (completes + partials)",
    x = "Selection rate", y = NULL
  )
```

### What ideal care looks like (q6)

Everyone gets q6 regardless of pathway (choose up to three), so it is the
question with the widest denominator — and likely a centerpiece of the
narrative analysis.

```{r}
#| label: ideal
#| fig-height: 5
responses |>
  filter(source == "survey", question == "q6", option != "other_text", !is.na(selected)) |>
  summarise(n_selected = sum(selected), n_shown = n(), .by = option_text) |>
  mutate(rate = n_selected / n_shown) |>
  ggplot(aes(rate, fct_reorder(option_text, rate))) +
  geom_col() +
  geom_text(aes(label = str_glue("{n_selected} of {n_shown}")), hjust = -0.2) +
  scale_x_continuous(labels = label_percent(), expand = expansion(mult = c(0, .15))) +
  labs(
    title = "What would ideal care support look like?",
    subtitle = "All pathways, survey respondents shown the question (choose up to three)",
    x = "Selection rate", y = NULL
  )
```

### Respondents' own words

Open-form "Other" answers are captured verbatim in `*_other_text` columns —
an early taste of the qualitative corpus before the dedicated open-text
questions arrive. Counts by question:

```{r}
#| label: other-text
survey |>
  select(ends_with("_other_text")) |>
  summarise(across(everything(), \(x) sum(!is.na(x)))) |>
  pivot_longer(everything(), names_to = "question", values_to = "responses") |>
  mutate(question = str_remove(question, "_other_text")) |>
  filter(responses > 0) |>
  arrange(desc(responses)) |>
  knitr::kable()
```

### Doorstep challenges (canvassing screener)

The canvasser-recorded quick challenges use short labels and a broader
denominator (everyone who engaged, not just full-question takers).

```{r}
#| label: doorstep
doorstep <- responses |>
  filter(question == "doorstep", !is.na(selected)) |>
  summarise(n_selected = sum(selected), n_shown = n(), .by = option_text)

ggplot(doorstep, aes(n_selected, fct_reorder(option_text, n_selected))) +
  geom_col() +
  geom_text(aes(label = n_selected), hjust = -0.5) +
  scale_x_continuous(expand = expansion(mult = c(0, .1)), breaks = breaks_width(1)) +
  labs(
    title = "Challenges mentioned at the door",
    subtitle = str_glue("Among {unique(doorstep$n_shown)} engaged contacts"),
    x = "Mentions", y = NULL
  )
```

## Survey demographics

The survey (not canvassing) carries demographics: household size and income,
age, gender, race/ethnicity, disability, political views, plus respondent ZIP.
These are the levers for the project's diversity targets and for subgroup
analysis once counts support it.

```{r}
#| label: demographics
survey |>
  select(age, hh_size, hh_income, disability, political) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "value") |>
  count(variable, value) |>
  knitr::kable()
```

## What analysis this supports

As monthly deliveries accumulate, this same structure supports:

- **Theme prevalence** — selection rates for challenges, supports, and ideal
  care, with uncertainty, overall and over time
- **Pathway comparisons** — do current, past, and future caregivers name
  different challenges and ideal supports?
- **Mode comparisons** — the shared battery lets survey and canvassing
  responses be pooled or contrasted
- **Geographic variation** — respondent ZIP (survey) and canvasser ZIP roll up
  to county/state for maps and for comparison against the state care-policy
  grades from Research Objective 1
- **Subgroup analysis** — demographic breakdowns of every rate above, gated on
  minimum cell sizes
- **Text analysis** — the open-text questions (q5, q7) are not in the test
  exports yet; when they arrive they feed the NLP pipeline in `scripts/nlp/`
